
%amslatex file
%\documentclass[amssymb,12pt,amscd,leqno]{amsart}
\documentclass[12pt,leqno]{amsart}
\usepackage[dvips]{graphics}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{color}
%\newenvironment{beweis}{\begin{proof}[Beweis]}{\end{proof}}
%\usepackage{epsfig}
%\typein[\answer]{To run syntax check, enter `S';
% otherwise, just press return:}
% \let\SANSWER\syntaxonly     \let\sANSWER\syntaxonly
% \csname\answer ANSWER\endcsname
\newenvironment{pf}{\proof[\proofname]}{\endproof}

\numberwithin{equation}{section}

%\voffset=-.7in
%\hoffset=-.7in
%\setlength{\textwidth}{7in}

%\setlength{\textheight}{8.5in}

\newtheorem{thm}{THEOREM}[section]
\newtheorem{thmmain}[thm]{\tref{thm-main}}%[section]
\renewcommand{\thethmmain}{}
\newtheorem{lem}[thm]{Lemma}
%\newtheorem{defn}[thm]{DEFINITION}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{PROPOSITION}
\newtheorem{conj}[thm]{CONJECTURE}
\newtheorem{quest}[thm]{PROBLEM}
\newtheorem{fak}[thm]{Fakt}

\newtheorem{hyp}[thm]{HYPOTHESES}
\newtheorem{ft}[thm]{FAKT}
\newtheorem{satz}[thm]{SATZ}
 \theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}%[section]
%\newtheorem*{proof}{Beweis}
\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem{ex}[thm]{Example}

\newcommand{\tref}[1]{Theorem~\ref{#1}}
\newcommand{\cref}[1]{Corollary~\ref{#1}}
\newcommand{\pref}[1]{Proposition~\ref{#1}}
\newcommand{\rref}[1]{Remark~\ref{#1}}
\newcommand{\dref}[1]{Definition~\ref{#1}}
\newcommand{\secref}[1]{\S\ref{#1}}
\newcommand{\lref}[1]{Lemma~\ref{#1}}
\newcommand{\exref}[1]{Example~\ref{#1}}
\newcommand{\sref}[1]{Satz~\ref{#1}}
\newcommand{\fref}[1]{Fakt~\ref{#1}}
\newcommand{\Ant}{\mathrm{Ant}}
\newcommand{\diam}{\mathrm{diam}}
\newcommand{\Pol}{\mathrm{Pol}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\pt}{\mathrm{pt}}
\newcommand{\rad}{\mathrm{rad}}
\newcommand{\scal}{\mathrm{scal}}
\newcommand{\Rad}{\mathrm{Rad}}
\newcommand{\Sub}{\mathrm{Sub}}
\newcommand{\Iso}{\mathrm{Iso}}
\newcommand{\Rg}{\mathrm{Rg}}
\newcommand{\Fix}{\mathrm{Fix}}
\newcommand{\R}{\mathbb{R}}

\def\X{{X^\varepsilon}}
\def\di{{\bold div}\,}

\def\RR{\mathbb R}
\def\co{\colon\thinspace}
\def\eps{\varepsilon}
\def\BAD{\mathop{\rm BAD}\nolimits}%


\def\:{\colon}
\newcommand*{\set}[2]{\left\{\,\left.{#1}\vphantom{#2}\,\right|\,{#2}\,\right\}}
\def\emptyset{\varnothing}
\newcommand*{\GHto}{\mathbin{\begin{picture}(16,3.5)
\put(-1,1){$\longrightarrow$}
\put(13,-3){\llap{\text{\sf\tiny GH}}}
\end{picture}}}


\newcommand{\curv}{\mathop{\rm curv}}
%MY STYLE
\def\parit#1{\medskip\noindent{\it #1}}
\def\parbf#1{\medskip\noindent{\bf #1}}
\def\qeds{\qed\par\medskip}
\def\qedsf{\vskip-6mm\qeds}


\pagestyle{plain}

\begin{document}
%\tableofcontents
\pagebreak
%\bibliographystyle{alpha}

%\pagenumbering{roman}

\title{Metric-measure boundary and geodesic flow  on Alexandrov spaces}
\thanks{\it 2010 Mathematics Subject classification.\rm\ Primary
53C20, 52A15, 53C23. Keywords: Alexandrov spaces, convex hypersurfaces, geodesic flow, Liouville measure}\

\author{Vitali Kapovitch, Alexander Lytchak and Anton Petrunin}
%\address{Mathematisches Institut\\ Universit\"at Bonn\\
%Wegelerstrasse 10, 53115 Bonn, Germany\\}
%\email{lytchak\@@math.uni-bonn.de}


%\subjclass{53C20,  52B99}
%\footnotetext[1]{}


%\keywords{Semi-convex functions, Alexandrov spaces, differentials}
%Spherical building, Euclidean building, non-positive curvature}

%\thanks{to Professor Ballmann}

%\date{\today}
%\date{August, 1997}

\begin{abstract}
 We relate the existence of many infinite geodesics on Alexandrov spaces to a statement about the average growth of volumes of balls. We deduce that the geodesic flow exists and preserves the Liouville measure in several important cases. The developed analytic tool has close ties  to integral geometry.
\end{abstract}


\maketitle
\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
\pagenumbering{arabic}

%\tableofcontents

\section{Introduction}
\subsection{Motivation and application}
 The following structural question in    the theory of Alexandrov spaces was formulated in a slightly different way in \cite{PP} and remained open since then.
\begin{itemize}
%\item Is the boundary of an Alexandrov space an Alexandrov space?
\item Are there  ``many'' infinite geodesics on any Alexandrov space without boundary?
\end{itemize}
%This question has been formulated in a slightly different way in \cite{PP} and remained open since then.
We address this question and obtain an affirmative answer in several cases, by
%In this paper we address the latter question, and obtain an affirmative answer in several cases.
developing  a new analytic tool. This tool, the Taylor expansion of the average volume growth, has a formal similarity with  classical integral geometry
and might  be interesting in its own right beyond the realm of Alexandrov geometry.

In particular, we prove the existence of such  infinite geodesics in the most classical examples of non-smooth Alexandrov spaces:

\begin{thm} \label{thmfirst}
Let $X$ be the boundary of a convex body in $\mathbb R^{n+1}$.
Then almost any direction in the tangent bundle $TX$
is the starting direction of a unique  infinite geodesic on $X$.
The geodesic flow is defined  almost everywhere and  preserves the Liouville measure.
\end{thm}

 Even  the existence of a \emph{single} infinite geodesic has apparently  not been known,  even in the case $n=2$,   \cite{Zam-quest}. Our result might  appear a bit surprising,
since on \emph{most} convex surfaces {\emph most points in the sense of Baire categories}   are not inner points of any geodesic,  due to \cite{Zam-inv}.
 %has second Baire cathegory}\footnote{\color{red}I think this formulation is slightly less confusing} due to \cite{Zam-inv}.


 \subsection{The geodesic flow and mm-boundary}
On a smooth manifold with boundary the geodesic flow is not defined for all times and the amount of geodesics terminating at the boundary in a given time interval depends on the size  of this boundary.
In the following definition we capture the size of the boundary by estimating the average volumes of small balls and their deviations from the corresponding  volumes in the Euclidean space.

Let $(X,d)$ be a locally compact separable metric space and let the Hausdorff dimension $n$ of $X$ be a natural number. Let $\mu$ be   a  Radon  measure on $X$ and assume that closed bounded subsets
of $X$ have finite $\mu$-measure.
For $x\in X$ and $r>0$
denote by $B_r (x)$ the open metric ball of radius $r$ around  $x$. Consider  the Borel function $b_r:X\to [0,\infty )$,
\begin{equation}
 b_r(x):=\mu (B_r (x)) \; .
 \end{equation}

We define the \emph{deviation measure} $v_r$ of $X$ as a signed Radon measure $v_r$ on $X$, absolutely continuous with respect to $\mu$,   given by
\begin{equation} \label{eq:first}
 v_r  = (1 - \frac {b_r} { \omega _n r^n} )\cdot \mu \, .
\end{equation}
Here $\omega _n$ is the volume of the $n$-dimensional unit Euclidean ball.
A {\emph signed Radon measure}, here and below, is  an element of the space $\mathcal M(X)$ dual to the space of compactly supported continuous functions $C_c (X)$.
We  consider $\mathcal M(X)$ with the topology of weak convergence. { Recall that a subset $\mathcal F \subset \mathcal M(X)$ is relatively compact if and only if 
it is \emph{uniformly bounded}, thus if for any compact subset $K\subset X$ the values $\nu  (K), \nu \in \mathcal F$ are uniformly bounded.}

In all examples and applications below,
 %$n$  is the Hausdorff dimension of $X$ and
 $\mu$ is the $n$-dimensional Hausdorff measure $\mathcal H^n =\mathcal H^n _X$ on $X$, always normalized so that it coincides with the Lebesgue measure on $\R^n$.


The curve $r\mapsto  v_r \in \mathcal M(X)$ controls in a very rough integral sense the deviation of $X$ from $\R^n$.
The Taylor coefficients of $v_r$ at $0$ (if they are defined) can be interpreted as integral curvatures, in analogy with integral geometry. The following  example is well known.



    \begin{ex} \label{smoothscal}
 Let $X^n$ be a  smooth Riemannian manifold with  Riemannian volume $\mu =\mathcal H^n$.  Then, for $r\to 0$,
  the measures $v_r /r^2$ converge in $\mathcal M(X)$ to $\frac 1 {6(n+2)}\cdot scal \cdot \mu$, where
  $scal$ denotes the scalar curvature of $X$.
\end{ex}



The next example can be obtained by computations in local coordinates. Since it is not needed in the sequel, we omit
  the details.




  \begin{ex} \label{mainex}
Let $X$ be a smooth $n$-dimensional Riemannian manifold with boundary $\partial X$.
Then, for $r\to 0$, the measures   $v_r/r$  converge in $\mathcal M(X)$ to
$c_n \cdot \mathcal H^{n-1} _{\partial X}$, for some constant $c_n >0$ depending only on $n$.
\end{ex}

%Here and below we consider the space of signed measures $\mathcal M (X)$ with the topology of weak convergence.
This example  suggests  to view the first Taylor coefficient of $v_r$   as the ``boundary'' of the metric-measure space $(X, d, \mu)$,
 which we make more precise by the following central definition of this paper.
% In the next definition, we say that a family of measures on a locally compact space $X$
%  is locally bounded if for any compact subspace $K$ of $X$, the restrictions of the family to $K$ are bounded.

\begin{defn} \label{def:first}
Let $(X,d,\mu)$ be a metric measure space as above.  Let   $v_r$ be the deviation measure of $X$,  as in \eqref{eq:first}.
%and let  the Radon measure $|v_r|$ be the total variation of $v_r$.
We say that $X$ has   locally finite metric-measure boundary, abbreviated as  mm-boundary,
if the  family of signed Radon measures
\[\{\, \frac { v_r }  {r } \,;\,  0<r\leq  1 \,\}\]
is uniformly bounded.
% we say that $X$ has   locally finite mm-boundary ({\color{red} or  metric-measure boundary}).
If  $\lim _{r\to 0} v_r  /r =\nu $ in $\mathcal M(X)$, we call $\nu$ the mm-boundary of $X$.  If $\nu =0$ we say that $X$ has trivial mm-boundary.
\end{defn}


We refer to Subsection \ref{subsec:example}  and Section \ref{sec:final}  below for a discussion of  examples and questions, and
 state now our central  result connecting  mm-boundaries to the existence  of infinite geodesics in Alexandrov spaces:

 \begin{thm} \label{thmmain}
 Let $X$ be an Alexandrov space. If $X$ has trivial mm-boundary, then  almost each direction of the tangent
 bundle $TX$ is the starting direction of  an infinite geodesic. In this case, the geodesic flow preserves the Liouville measure on $TX$.
 \end{thm}

In different  settings, geodesic flow  on singular spaces have been investigated in \cite{BallmannBrin} and \cite{Bamler}.

\subsection{Size of the mm-boundary in Alexandrov spaces}
The next theorem shows that, similarly to \exref{mainex},  the topological boundary is closely related to the mm-boundary in Alexandrov spaces.

\begin{thm} \label{alexandrovthm}
Let $X^n$ be an  $n$-dimensional Alexandrov space. Then $X$ has locally finite mm-boundary.
If  $ \nu= \lim     \frac{v_{s_j}}{s_j}$,  for  a sequence  $s_j \to 0 $,  then $\nu$ is a Radon measure and
 the following holds true.
\begin{enumerate}
\item\label{full-measure-zero-nu}There is a Borel set $A_0$ with $\mathcal H^n (X\setminus A_0) = \nu (A_0)=0$.
\item\label{bry-nu} If the topological boundary $\partial X$ is non-empty then $\nu \geq c \cdot \mathcal H^{n-1} _{\partial X}$,
for a positive constant $c$ depending only on $n$.
\item \label{n-1-nu} If the topological boundary $\partial X$ is empty then $\nu (A)=0$, for any Borel subset $A\subset X$ with  $\mathcal H^{n-1} (A)<\infty$.
\end{enumerate}
\end{thm}

 We strongly believe that an Alexandrov space with empty topological boundary $\partial X$  has a trivial mm-boundary,
which would solve the question about the existence of infinite geodesics.
We could prove this conjecture in two special cases:



\begin{thm} \label{hypersurface}
Let $X^n$ be a convex hypersurface in $\R^{n+1}$  or let $X$ be a two-dimensional  Alexandrov space without boundary. Then $X$ has trivial mm-boundary.
\end{thm}


In combination with \tref{thmmain} this proves \tref{thmfirst}.
The two-dimensional case
will follow from a much stronger result discussed in the next subsection.
  Another proof of  the  two-dimensional case could be derived   from  the statement about convex hypersurfaces and  Alexandrov's embedding theorems.
% if $X$
%is non-negatively curved. This approach can be extended to cover all
%\tref{thmfirst} for the hyperbolic space,
%since by  Alexandrov's embedding theorem, any two-dimensional Alexandrov space without boundary
%is locally isometric to a convex surface in $\mathbb H^3$.\footnote{ does this hold for $\mathbb H^3$ of curvature $-1$?Vitya}



\subsection{mm-curvature} \label{subsec:curv}
Motivated by Example \ref{smoothscal} one can naively hope that the second Taylor coefficient at $0$ of the
map $v_r\colon r\mapsto \mathcal M(X)$ describes the scalar curvature of the space.
%This again has  natural analogy with the classical integral geometry.



\begin{defn}
 Let $X,v_r$ be as  in Definition \ref{def:first} above. If the family $v_r /r^2, r\leq 1$ is uniformly  bounded then we say that $X$ has locally finite  mm-curvature.
 % For any compact subspace $K$ of $X$, we call $\limsup _{r\to 0} \|v_r /r\| _K$ the bound of the scalar curvature  in $K$.
  If the measures $v_r /r^2$  converge to a measure $\nu$, we call $\nu$ the mm-curvature of $X$.
\end{defn}

%If, in the definition above, the measures $v_r /r^2$ locally converge to a measure $\nu$, we call $\nu$ the mm-curvature of $X$.
Clearly, local finiteness of mm-curvature as defined above implies that the mm-boundary is trivial.
Thus, the following result proves \tref{hypersurface} in the $2$-dimensional case.


\begin{thm} \label{intsurface}
 Let $X$ be a 2-dimensional Alexandrov space without boundary.
 Then $X$ has locally finite mm-curvature.
% Moreover, it is bounded by a linear function of $|\omega |$, at least for
% if $|\omega|$ is small enough.
\end{thm}


This finiteness result holds true in the much greater generality of surfaces  with bounded integral curvature in the sense of Alexandrov--Zallgaler--Reshetnyak \cite{Reshetnyak-GeomIV}, \cite{AZ}, see Section \ref{sec:surface} below.

Note, however, that the mm-curvature
in \tref{intsurface} does not need to coincide with the ``curvature measure'' as  defined in \cite{AZ},   even in the case of a cone, cf. Example \ref{ex:cone}.  In particular, this shows that the mm-curvatures in $2$-dimensional Alexandrov spaces are not stable under Gromov--Hausdroff convergence.


\begin{rem}
N. Lebedeva and the third named author have found in \cite{LP}  a ``scalar curvature measure'' on all smoothable Alexandrov spaces.
% which has the property of being stable
%under non-collapsed Hausdroff convergence.
There is a hope, supported by our proof of  \tref{intsurface}, that a better understanding of this ``stable curvature measure'' will
lead to some control of the   mm-boundary and mm-curvature discussed here.   
\end{rem}



\subsection{Analogy with the Killing--Lipschitz curvatures and the proof of \tref{thmmain}}
For a compact convex body or compact smooth submanifold $M$ in $\R^n$, one classically considers
the  volumes $w(r)=\mathcal H^n (B_r(M))$ of distance tubes $B_r(M)$ around $M$.
It turns out that $w(r)$ is a polynomial, at least for small $r$, and  the coefficients of
$w$, called the   Killing--Lipschitz curvatures of $M$, are given by  integrals over some intrinsically defined curvature
 terms.  Moreover, these coefficients can be localized and considered as measures on $M$. We refer  to \cite{Alesker} for a short account of the theory, connection of the theory with 
\cite{LP} and further hypothetical relations with the theory of Alexandrov spaces. 

To make the formal similarity with our approach  to mm-boundary and mm-curvature more transparent, we observe that
(at least for a smooth $n$-dimensional manifold $M$) the number   $\int _M \mathcal H^n  (B_r(x))\, \cdot  d\mathcal H^n (x)$  can be interpreted as the $\mathcal H^{2n}$-measures
of the distance tubes $B_{\frac r {\sqrt 2} } (\Delta )$ around the diagonal $\Delta $ in the Cartesian product $M\times M$.

The natural interpretation of the tangent bundle of $M$ as the normal  bundle of the diagonal $\Delta$ in $M\times M$ gives a connection between
the measure theoretical properties of the tubes around $M$ and the dynamical properties of the geodesic flow.
We clarify this abstract statements by explaining the main idea of  our  proof of \tref{thmmain} in the  well-known case of a complete smooth Riemannian manifold $X=M$.
In this case the existence of  geodesics is trivial.
Thus, we just sketch a new proof  of the  classical fact   that the geodesic flow $\phi$  preserves the Liouville measure $\mathcal M$ on $TM$.
This proof is stable and rough enough to be transferred to the singular situation,
% but
%The idea can be fastest explained  in the case where $X$ is a smooth Riemannian manifold $M$,
%where it provides a metric-measure  proof of the most classical  statement that the geodesic flow $\phi_t\:TM\to TM$ preserves the Liouville  measure $\mathcal M$ on $TM$. It suffices to consider the case $t=1$
%and to prove the statement on arbitrary small neighborhoods of the $0$-section.
%For any natural $n>0$, we can write $\phi _t =\phi _{\frac t n} ^n$, thus it is sufficient  to understand this measure preservance for small $t$, or even infinitesimally.

Let  $\phi_t\: TM\to TM$ be the geodesic flow for time $t$.
Denote by $\pi\: TM\to M$ the tangent bundle of $M$. Define the map $E\:TM\to M\times M$  by
\[E(v)=(\pi(v), \pi (\phi _1 (v)).\]
By construction, $E(-\phi_1(v))= J(E(v))$,
where $J$ is the involution of $M\times M$ which switches the coordinates.
Since $J$ preserves $\mathcal H^{2n}$ on $M\times M$ and $v\to -v$ preserves the Liouville measure $\mathcal M$ on $TM$,
the statement that $\phi$ is measure preserving hinges upon the smallness of measure-distortion of the map $E \:(TM,\mathcal M) \to (M\times M, \mathcal H^{2n})$ close to the $0$-section.
In the present case of a Riemannian manifold,
this property of $\phi_1$ is expressed by  the fact that the differential of $E$ is the identity (after suitable identifications).
In the general case of Alexandrov spaces, we similarly observe that the ``infinitesimal'' deviation (via the canonical map $E$) between $J$ being measure preserving (which we know) and  $\phi _1$ being measure preserving (which is what we want to show)
is expressed precisely as the triviality of  the mm-boundary.






  %In fact, we  restrict here  to the case of a smooth Riemannian manifold $M$ and provide a metric explanation for the invariance of the Liouville measure under the geodesic flow $\phi _t$.  Note that $   We discretize  of the geodesic flow and think of it as the limiting process, for $t\to 0$ of going
%in the classical case of a smooth Riemannian manifold $M$.





\subsection{Stability and relation with quasi-geodesics}
Many Alexandrov spaces, for instance all convex hypersurfaces, appear naturally  as Gromov--Hausdorff limits of smooth Riemannian manifolds. However, the properties of the geodesic flow, mm-boundaries and mm-curvature are  unstable under limit operations, cf. the discussion at the end of Subsection \ref{subsec:curv}. Thus, there is no hope to deduce \tref{intsurface}, \tref{hypersurface} or \tref{thmfirst} by a direct limiting argument.

For instance,  being a geodesic is a local notion, not preserved under limits.  However, any limit of geodesics  in  a non-collapsed
limit  of Alexandrov spaces  is a so called quasi-geodesic, a  curve sharing many properties with geodesics.  Quasi-geodesics have been investigated  in \cite{PP}, cf. \cite{Petsemi} and the references therein. It was shown that any direction is the starting direction of an infinite quasi-geodesic.
One motivation for the present paper was an attempt to prove Liouville's theorem for the ``quasi-geodesic flow'', see Subsection
\ref{subsec:quasi} below.
%In Alexandrov spaces  geodesics often cannot be extended as local geodesics beyond some point.
%This sudden ``death''  of geodesics  causes  serious technical difficulties.
%To deal with this problem  so called \emph{quasi-geodesics} (certain generalized geodesic)
%were introduced
%and it was proved
%that any direction is the starting direction of an infinite quasi-geodesic;
%see \cite{Petsemi}, \cite{PP} and the references therein.

%For most application the existence of quasi-geodesics is sufficient.%
%Quasi-geodesics are in many senses as good as geodesics, and even better in the sense of their stability.
%However, quasi-geodesics in a given direction are not uniquely defined, they can branch.
%One motivation for the present paper was an attempt to prove Liouville's theorem for the ``quasi-geodesic flow''.

%Unlike local geodesics, quasi-geodesics are defined by a global metric condition. Therefore they enjoy much better stability properties
%and can often be found by limiting arguments.
%We note that while many Alexandrov spaces appear naturally  as limits of smooth Riemannian manifolds, the properties of the geodesic flow, %mm-boundaries and mm-curvature are unstable under
%convergence.
%For instance, there is  no chance to deduce our main results about convex hypersurfaces by approximating them by smooth or polyhedral ones.



\subsection{Examples} \label{subsec:example}
The estimates of the   mm-boundary and mm-curvature are quite involved even in  quite  simple situations.
We collect a few instructive examples. %indicating relations with curvature measures in integral geometry.
The examples  are not  needed in the sequel and we omit the somewhat tedious computations.   Examples \ref{ex:cone}, \ref{ex:secondlast} and \ref{ex:last} should be compared with
\cite{Bernig-CAT} and \cite{Bernig-Alex} revealing further natural connections to the theory of Lipschitz--Killing curvature on singular subsets of the Eucldiean space.
%\footnote{A.: How instructive are the examples? All of them?{\color{red} I think the  first one can be dropped. The last 3 I would keep. Not sure about keeping the second one without a proof. Vitya}
% A.:   But one also might add the following one, together with the comment, that $\mu =\mathcal H^n$ is the only measure for which our considerations are of some interest.  Maybe, it is better not to say it.

\begin{ex} Let $X$ be a Lipschitz manifold and $\mu=\mathcal H^n$.
Then $\lim _{r\to 0} v_r =0$.  Moreover, $\mathcal H^n$ is the only measure with this property.
\end{ex}












\begin{ex}
Let $X$ be a Riemannian manifold with a Lipschitz continuous metric.
Then $X$ has  trivial mm-boundary.
\end{ex}


 \begin{ex}
  If $X$ is a manifold with  two-sided bounded curvature in the sense of Alexandrov then its
  mm-curvature   is  well-defined and  absolutely continuous with respect to the Hausdorff measure.
 \end{ex}




\begin{ex} \label{ex:cone}
 Let $X$ be the flat cone $CS_{\rho}$ over a circle of length $\rho$.
  Then $X$ has finite mm-curvature. The mm-curvature is a Dirac measure of mass $m(\rho)$ concentrated in the origin.
  If $\alpha =2\pi-\rho$ denotes the defect of $X$ then the classical curvature measure is the Dirac measure of total mass
  $\frac \alpha 6$.  On the other hand $m(\rho)= \frac \alpha 6 + f(\alpha)$, where
  $f(\alpha)$ is a non-zero function which has at the origin the form $f(\alpha)= O(\alpha ^3)$.
   \end{ex}




 \begin{ex}  \label{ex:secondlast}
 Let $X$ be a finite $n$-dimensional simplicial complex with an intrinsic metric $d$.
 Assume that the restriction of $d$ to each simplex is given by a (sufficiently) smooth Riemannian metric.
 Then $X$ has a finite boundary $\nu$, concentrated on  the $(n-1)$-skeleton $X^{n-1}$.
 %Moreover, the restriction of $\nu $ to
 %$X^{n-1}$ is given by $\nu= c_n \cdot f \cdot \mathcal H ^{n-1}$.
 %Here $c_n$ is the positive constant from Example  \ref{mainex} and  $f$ is the locally constant function $f(p)=(2-m)\cdot m$, where $m$ is the number of top-dimensional simplices
 %adjacent to a point $p\in X^{n-1}$.
 \end{ex}


 \begin{ex}  \label{ex:last}
 Assume that $X$   as in the last example   is a pseudo-manifold.
Then $X$ has finite mm-curvature.
%\footnote{A.: I have killed the long description of the mm-curvature?  Would it be better to keep it? {\color{red} Either way is fine with me. Vitya}}
% given by $\mu = \mu _1 + \mu _2 + \mu _3$.  Here $\mu_1$ is  the usual scalar
%    curvature on maximal faces. The second term  $\mu _2$ lives on $ X^{n-1}$, and is given by
%    $q_n \cdot m(p) \cdot \mathcal H^{n-1}$.
%    Here $m(p)$ is the sum of mean curvatures and $q_n$ is a constant depending only on the dimension.
%    Finally, $\mu_3$ lives on the $(n-2)$-skeletton  $X^{n-2}$ and is given by $w_n \cdot c(p) \cdot \mathcal H^{n-2}$.
%     Here $w_n$ is a constant depending on the dimension and $c(p)$ depends only on the link at $p$.
 \end{ex}



\subsection{Structure of the paper}
After preliminaries collected in Section \ref{sec:prelim}, we prove \tref{thmmain} in Section \ref{sec:Liou} along the lines sketched above.
In Sections \ref{sec:surface}, \ref{sec:hyper}  and \ref{sec:Alex} we prove  the remaining Theorems \ref{intsurface}, \ref{hypersurface}, \ref{alexandrovthm}, respectively. The proofs of these theorems all rely on a decomposition of the space into a regular and a singular part, with
a quantitative estimate of the size of the singular part. Finally, on the regular part we estimate the mm-curvature and mm-boundary through other
canonical measures  on these spaces. In the case of surfaces, this comparison measure is the classical curvature measure,  in the case of convex hypersurfaces, this comparison measure is the mean curvature.  Finally, in the case of a general  Alexandrov space, the comparison is given by the derivative of the metric tensor  expressed in
%canonical\footnote{ \color{red}I am not sure  how appropriate it is to use the word "canonical" here.}
DC-coordinates, \cite{Per-DC}. The needed control of the ball growth
in terms of these measures is given by a theorem of M. Bonk  and U. Lang in the case of surfaces  and follows from classical convex geometry in the case of hypersurfaces. The analytical comparison result  needed for Alexandrov spaces is established in Section \ref{sec-BV-estimate}. In the final
Section \ref{sec:final} we collect a number of comments and  open questions which naturally arose during the work on this paper.




\subsection{Acknowledgments}  The authors are grateful for helpful conversations  and comments to  Semyon Alesker, Andreas Bernig, Aaron Naber, Koichi Nagano, Richard Bamler.
The first author was supported in part by a Discovery grant from NSERC and by  a Simons Fellowship from the Simons foundation (award 390117). The second author was supported in part by the DFG grants  SFB TRR 191 and SPP 2026.


\section{Preliminaries} \label{sec:prelim}
\subsection{Metric spaces}
We refer to \cite{BBI01} for basics on metric spaces.
The distance between points $x,y$ in a metric space $X$ will be denoted by $d(x,y)$.
% or $|xy|_X$ or just $|xy|$.
By $B_r(x)$ we will denote the open metric ball of radius $r$ around a point $x$. For $A\subset X$
we denote by $B_r (A)$ the open $r$-neighborhood $B_r (A) =\cup _{x\in A} B_r (x)$.


A \emph {minimizing geodesic} $\gamma$ in a metric space $X$ is  a map $\gamma  : \mathbb I\to X$ defined on  an interval $\mathbb I$  such that for some number $\lambda \geq 0$  and all $t,s\in \mathbb I$
$$ d(\gamma (t),\gamma (s)) =\lambda \cdot |t-s| \;.$$
 In particular, we allow $\gamma$ to have any constant velocity $\lambda \geq 0$.
 % On the other hand, geodesics
%as defined here will always be globally ``minimizing geodesics'' in the usual notations of Riemannian geometry.
A \emph{geodesic} is a curve $\gamma\: \mathbb I\to X$ such that its restriction to a small neighborhood of any point in $\mathbb I$ is  a minimizing geodesic. Note that a  geodesic is a curve of constant velocity.

\subsection{Metric measure spaces} We refer to \cite{Federer} and \cite{Evans} for basics on measure theory needed here.

Let $X$ be a locally compact separable metric space. A \emph{Radon measure} on $X$ is a measure on $X$ for which all compact subsets are measurable and have finite measure.
Any Radon measure is canonically contained in the dual space $\mathcal M(X)$ of the topological vector space $\mathcal C_c (X)$ of compactly supported continuous functions on $X$.
All elements in $\mathcal M(X)$ are called \emph{signed Radon measures}. For any signed Radon measure $\mu$ there exist unique Radon measures $\mu ^{\pm}$ such that $\mu =\mu ^+- \mu ^-$.
In such a situation, we will denote by $|\mu |$ the Radon measure $\mu ^+ +\mu ^-$ and call it the \emph{total variation} of $\mu$.


%For a Radon measure $\mu$ on $X$, the $n$-dimensional density of $\mu$ at a point $x$ is defined by  $\limsup _{r\to 0} \frac {\mu (B_r(x))} {\omega _n r^n}$.

 A family   $\mathcal F$ of signed  Radon measures on $X$  is \emph{uniformly bounded} if  for any compact subset $K\subset X$ there exist a constant $C(K)>0$ such that $|\mu| (K) \leq C(K)$ for any $\mu \in \mathcal F$.
   Any uniformly bounded sequence of signed measures $\mu _i$  has a convergent subsequence.

  %Let $\nu _s, 1\geq s >0$ be a bounded family of Radon measures. For any sequence $s_j\to 0$, the sequence $\nu _{s_j}$ has a convergent
%subsequence. The set of all limits of all such subsequences is a bounded family of Radon measures and has a well defined least upper bound Radon measure $\mu$. In  this situation we will write $\mu =\limsup _{s\to 0} \nu _s$.

We will repeatedly use the following simple


\begin{lem} \label{lem:exchange}
Let $X$ be a metric space with two Radon measures $\mu $ and $\nu$. Let $r>0$ be arbitrary and let  $A\subset X$ be a Borel subset. Then
 $$\int _A \mu (B_r(x)) \cdot d\nu (x) \leq \int _{B_r (A)}  \nu (B_r (x)) \cdot d\mu (x) \; .$$
\end{lem}

\begin{proof} Due to  Fubini's theorem
the left  hand side is the volume of  $$S=\{(y,x) \in X\times X \;| \;  y \in A,\, d(y,x) <r \}$$ with respect to the product measure $\nu \otimes \mu$.
And on the right  hand side of the inequality is the volume of the larger set $$T=\{(y,x) \in  X\times X \; | \; x\in B_r(A), \, d(y,x)<r \}$$ with respect to the same measure.
\end{proof}

\subsection{Alexandrov spaces} \label{subsec:Alex}
We are assuming that the reader is familiar with basic theory of Alexandrov spaces and  refer to ~\cite{BGP} as a good introduction to the subject.
Throughout this paper, an Alexandrov space will denote a complete, locally compact, geodesic metric space of finite Hausdorff dimension and of curvature bounded from below by some $\kappa \in \R$.
%All Alexandrov spaces considered in this paper are finite dimensional.
 For Alexandrov spaces, an upper index will indicate the Hausdorff dimension; that is, $X^n$ denotes an $n$-dimensional Alexandrov space, equipped with the $n$-dimensional Hausdorff measure $\mathcal H^n$.
% In this situation we will sometimes denote
% $\mathcal H^n$ by $\vol_n$.

The set of starting directions of geodesics starting at a given point $x\in X$ carries a natural metric, whose completion is the {\emph tangent space} $T_x=T_xX$
of $X$ at the point $x$.  It is again an $n$-dimensional Alexandrov space of non-negative curvature. Moreover, it is the Euclidean cone
over the space $\Sigma _x$ of unit directions. The Euclidean cone structure defines multiplications by positive scalars $\lambda \geq 0$ on $T_xX$. The origin
of the cone $T_xX$ is denoted by $0=0_x$. Elements of $T_xX$ are called {\emph tangent vectors} at $x$. For $v\in T_xX$ the {\emph norm} $|v|$ of $v$ is the distance of $v$ from the origin $0_x$.


 Geodesics in $X$ do not branch and any two geodesics identical  starting vectors coincide.
For  $x\in X$ the \emph{exponential map}  $\exp_x $ is defined as follows.  Let $D_x$ denote the set of all vectors $v\in T_xX$
 for which there exists an (always unique)  \emph{minimizing geodesic}  $\gamma _v:[0,1] \to X$ with starting direction $v$.  The exponential map is defined on $D_x$ as
 $$\exp _x (v)= \gamma _v(1) \;.$$
%  Hence,
%for any $v\in T_xX$ there exists at most one minimizing geodesic $\gamma _v$  starting in $x$ in the direction of $v$ and defined on a maximal possible interval $\mathbb I =[0,t]$ with $t\in [0, \infty )$
%or $\mathbb I=[0,\infty)$.  Let $D_x$ denote the set of all vectors $v\in T_xX$ for which $\gamma _v(1)$ is defined.
%We set $\gamma _v (1)=\exp_x (v)$.
 For  any $r>0$, the map $\exp_x$ sends  $D_x\cap B_r (0_x)\subset T_xX$ surjectively onto $B_r(x)\subset X$.
 Moreover, for a constant $C=C(\kappa)\geq 0$  and all $r< \frac 1 C$ the map $\exp_x \:D_x \cap B_r(0_x) \to B_r(x)$ is  $(1+ C \cdot  r^2)$-Lipschitz continuous.



By the theorem of Bishop--Gromov, the volume $b_r (x)=\mathcal H^n (B_r(x))$ is bounded from above by the corresponding volume in the
space of constant curvature $\kappa$. In particular, $b_r (x) \leq \omega _n \cdot r^n + C \cdot r^{n+2}$ for all $r\leq \frac 1 C$, where
the constant $C$ can be chosen as before.  Thus, the deviation measures  $v_r $ from \eqref{eq:first} satisfy
$$v_r \geq -Cr^2 \mathcal H^n \; ,$$  for all sufficiently small $r$.
 Here and in the previous paragraph,  one can set $C=0$ if $\kappa \geq 0$.

% {\color{red} From now on we will assume that $k=0$ and $C=0$ in all proofs. This implies in particular that $v_r\ge 0$ and $\exp_x$ is $1$-Lipschitz. The proofs can be  easily adjusted to the case of general $k$ using the above estimate.}\footnote{{\color{red}I suggest we do all the proofs for $k=0$. Vitya}  It is fine with me. However, in the first 5 chapters it appears only here and  between euqation 3.4 and 3.5 in the proof of Theorem 3.5}


Denote by $X_{reg}$ the set of all points $x\in X$ with $T_xX$   isometric to $\R^n$.  The set $X_{reg}$ has full $\mathcal H^n$-measure in $X$.
Any inner point of any geodesic starting on $X_{reg}$ is contained in $X_{reg}$, \cite{Petparallel}.


The topological boundary $\partial X$ of $X$ can be defined as the closure of the  set of all points $x\in X$ with $T_xX$ isometric to a Euclidean half-space. Up to a subset of Hausdorff dimension $n-2$, $\partial X$ is an $(n-1)$-dimensional Lipschitz manifold.
%On the other hand, $X\setminus (X_{reg} \cup \partial X)$ has Hausdorff dimension at most $n-2$~\cite{BGP}.\footnote{\color{red}This is repeated in section 7. perhaps cut it either here or in section 7? Vitya}


\subsection{Volume and Bi-lipschitz maps}
It is easy to get an estimate on $v_r$  on parts which are Lipschitz manifolds and where one can control the biLipschitz constants.
Let $\mu =\mathcal H^n$ be a Radon measure on the metric space $X$.
Let $U\subset X$ and $V\subset \R^n$ be open and assume that there is an $(1+\delta)$-biLipschitz map
$f\:U \to V$.  %Assume $\delta  \leq \delta _n$, where $\delta_n$ is a sufficiently small constant depending on $n$.
 Let $A\subset U$ be given with
$B_{(1+\delta) \cdot  r} (A) \subset U$ and $B_{(1+\delta) \cdot r} (f(A)) \subset V$.  Then, for all $x\in A$,
\begin{equation} \label{eq:bilip}
(1+\delta)  ^{-2\cdot n} \leq \frac {b_r(x)} {\omega _n \cdot  r^n} \leq (1+\delta ) ^{2\cdot n} \; .
\end{equation}
Therefore, if $\delta $ is sufficiently small,  $|v_r | (A) \leq  3 n \cdot   \delta \cdot \mathcal H^n (A).$


%\begin{rem}  \footnote{A.: Tosha, do you know it is true?} For Alexandrov  spaces the opposite seem to hold (but I am not quite sure):
%If the volume of  a ball $B_r (x)$ satisfies $\omega _n r^n -\mathcal H^n (B_r (x)) \leq \delta r ^n$, then
%the ball $B_{\frac r 4} (x)$ is  $(1+C_n \delta )$-biLipschitz  to a Euclidean ball.
%\end{rem}





\section{Liouville measure and geodesics} \label{sec:Liou}
\subsection{Tangent bundle and Liouville measure} \label{subsec:tb}
Let $X$ be an $n$-dimensional Alexandrov space.
Denote by $TX$ the  disjoint union of the tangent spaces at different points,
\[TX=\bigsqcup_{x\in X} T_x X.\]
Let $\pi\:TX\to X$ be the canonical footpoint projection, so $\pi (T_xX)=\{x\}$ for any $x\in X$.
%Let $X^n$ be an $n$-dimensional Alexandrov space. Consider the tangent bundle of $X$
%  as the  disjoint union $TX=\cup T_x X$ of the tangent spaces at different points. Denote by $\pi\:TX\to X$ the canonical footpoint projection, $\pi (T_xX) =\{x\}$.	
For a subset $K\subset X$ denote by $TK$ the inverse image $\pi^{-1} (K)= \cup _{x\in K} T_xX$. Given $r>0$, denote by $T^r K$ the set of all vectors in $TK$
	of norm smaller than $r$.
% Let $X_{reg}\subset X$ be the set of all regular points and $TX_{reg} =\pi^{-1} (X_{reg}) \subset TX$.

 The  Riemannian structure on the set of regular points discussed in  \cite{Otsu-Shioya}  (cf.  \cite{Shioya}, \cite{Per-DC}) provides  $TX_{reg}$ with a canonical structure of a Euclidean vector bundle over $X_{reg}$. In this topology, for any sequence of geodesics $\gamma _i$ in $X_{reg}$ converging to a geodesic $\gamma$, the starting directions of $\gamma _i$ converge to the starting direction of $\gamma$.

 On the Euclidean vector bundle $TX_{reg}$ over $X_{reg}$ we have a canonical measure, which locally coincides with the product measure of $\mathcal H^n _X$ and the Lebesgue measures on the fibers. More precisely,
 it is the unique Borel measure $\mathcal M$ on $TX_{reg}$ such that for any  Borel set $A\subset TX_{reg}$
 $$\mathcal M(A)= \int _X \mathcal H^n(A \cap T_x X) \cdot d\mathcal H^n (x) \; .$$
 We extend $\mathcal M$ to a measure on $TX$ by setting $\mathcal M(TX\setminus TX_{reg})$ to be $0$.
By definition, a subset $A\subset TX$ is $\mathcal M$-measurable if and only if there exists  a Borel subset $A'\subset A\cap TX_{reg}$ such that
for $\mathcal H^n$-almost all $x\in X$ the intersection $(A\setminus A') \cap T_xX$ has $\mathcal H^n$-measure zero in $T_xX$.




 For any $\lambda >0$ the multiplication by $\lambda$  on  $T_xX$ sends $\mathcal M $ to $\lambda ^n \cdot \mathcal M$.
The canonical involution $I\:TX_{reg}\to TX_{reg}$, defined by $I(v)=-v$,
 preserves $\mathcal M$, since it preserves the Lebesgue measure in each tangent space.
%We extend $I$ to an arbitrary Borel map outside of $TX_{reg}$ (for instance, $I(v)=v$ for all $v\notin TX_{reg}$).


\subsection{Geodesic flow}
   In order to define the geodesic flow $\phi$ on a maximal subset   $\mathcal F$ of $TX \times \R $  we proceed as follows.
 For any  $v\in  T_xX$ we set $\phi _0(v)=v$.
 If no geodesic starts in the direction of $v$,
 the value $\phi _t(v)$ will not be defined for $t\neq 0$. If such a geodesic $\gamma_v$ exists, then $\gamma_v$   can be uniquely extended to a maximal possible half-open  interval $\gamma_v \:[0,a)\to X$.  For $t\geq a$
 the value $\phi _t(v)$ will not be defined. For $0<t<a$ we set $\phi _t(v)$ to be $\gamma _v ^+ (t) \in T_{\gamma _v(t)} X$, the starting direction of $\gamma_v\:[t,a) \to X$ at $\gamma_v (t)$.

 Whenever, the geodesic $\gamma _v\:[0,a)\to X$ extends to an (again uniquely defined, maximal)  geodesic   $\gamma_v\:(b,a) \to X$ for some $b<0$
 then we define $\phi _t(v)$ for $b<t<0$ to be $\gamma_v ^+ (t) $ as above.


For $\lambda >0$, for $t,s\in \R$  and $v\in  T_x X$ we have
 $$\phi _t(\lambda \cdot  v) :=\lambda \cdot \phi _{\lambda  \cdot t} (v) \; \;  \text{and}  \;  \;   \phi _{t+s} (v) =\phi _t ( \phi _s (v)) \; ,$$
 whenever the right hand side is defined.
% $$\phi _t ( \phi _s (v))= \phi _{t+s} (v) \; ,$$whenever the left hand side is defined for $t,s\in \R$.
The partial flow $\phi$ preserves the norm of tangent vectors.  Since inner point of geodesics starting in $X_{reg}$ are contained in $X_{reg}$, the set  $TX_{reg}$ is invariant under the  flow $\phi$.
%On $TX_{reg}$ we have
%$I\circ \phi_t = \phi _{-t} \circ I$, whenever both sides are defined.


By construction, the domain of the definition of the  geodesic flow almost includes the domain of the definition of the exponential map.
More precisely, consider the set
\[D=\bigcup_{x\in X} D_x\subset TX;\]
that is, the set of all vectors $v\in TX$ for which $\exp _{\pi(v)} (v)$ is defined.
Note that $\lambda\cdot D\subset D$ for any $0\leq \lambda \leq 1$.
Moreover, for all $v\in D$ and all $\lambda <1$
 the geodesic flow $\phi_1 (\lambda v)$ is defined and $\pi (\phi_1 (\lambda v))=  \exp _{\pi (v)} (\lambda v)$.


Thus,  for $\mathcal M$-almost all $v\in D$ we have the following
\begin{itemize}
\item $v\in TX_{reg}$,
\item $\phi _1(v) \in TX_{reg}$ is defined,
\item $w=-\phi_1 (v)\in D$ and
 \begin{equation} \label{eq:symm}
(\pi (w), \exp (w))=(\exp (v), \pi (v)) \in X\times X.
 \end{equation}
\end{itemize}


\subsection{Measurability} \label{subsec:measur}
In order to use measure-theoretic arguments, we need to ensure that the partial geodesic flow $\phi:\mathcal F\subset (TX\times \R) \to TX$ on its domain of definition $\mathcal F$
is measurable.

For any  geodesic $\gamma :\mathbb I\to X$ defined on a compact interval $\mathbb I$ there exists some $k>0$ such that  the restriction of $\gamma$ to any subinterval of length $\frac 1 k$ is
a (minimizing) geodesic. For bounded length of $\mathbb I$, bounded velocity of the curves and bounded $k$ any sequence of such  geodesics contained in a relatively compact set
contains a subsequence converging to a  geodesic. From this observation we can easily write  $\mathcal F\cap (TX_{reg} \times \R)$ as a countable union of subsets $\mathcal F_k$ closed in
$TX_{reg} \times \R$, such that the restriction $\phi: \mathcal F_k \to TX_{reg}$ is continuous.

Therefore,  the intersection $\mathcal F\cap (TX_{reg} \times \R )$ is a Borel subset  and the geodesic flow
$\phi :\mathcal F\cap (TX_{reg} \times \R)  \to TX_{reg}$ is a Borel-measurable map, see also Subsection \ref{subsec:quasi} below.








\subsection{Liouville property}
 Denote by $\mathcal G$ the set of all vectors $v\in TX $ such that $\phi _t(v)$ is defined for all $t\in \R$. By definition, $\mathcal G$ contains the $0$-section, it
 is invariant under multiplications by any $\lambda >0$  and it is invariant under the geodesic flow $\phi$. Moreover, $\mathcal G\cap TX_{reg}$ is invariant under the involution $I (v)=-v$.
  \begin{defn}
 We say that an Alexandrov space $X$ has the \emph{Liouville property} if  $TX\setminus \mathcal G$ has zero Liouville measure  and if
for any $t\in \R$ the geodesic flow $\phi _t \:\mathcal G\to \mathcal G$ preserves the Liouville measure.
 \end{defn}

 %  For a subset $K\subset X$ and a number $r<0$ we use the notation $T^ r K$
%to denote the set of all vectors in $\pi^{-1} (K)=\cup _{x\in K} T_xX$ of norm $< r$.
The Liouville property can be checked infinitesimally as follows.

\begin{lem} \label{infini}
An Alexandrov space $X$ \emph{does not} have the Liouville property if and only if there is a compact subset $K\subset X$, a positive number
$\epsilon$ and a sequence of positive numbers  $r_m \to 0$  with the following property.
 For every  $m$, there exists a Borel subset $A_m\subset T^{r_ m} K$ such that
 \begin{equation} \label{eq:m}
\epsilon \cdot r_m^{n +1} \leq \mathcal M (A_m) -\mathcal M (\phi _1 (A_m)) \;.
\end{equation}
\end{lem}
By an abuse of notations we denote in \eqref{eq:m}  and below by  $\phi_1 (A_m)$ the set of all $\phi _1(v), v\in A_m$, for which
$\phi _1(v)$ is defined.


\begin{proof}
If at least one $r_m$ with the above property exists, then $X$  does not have the Liouville property by  definition.

Assume on the other hand, that $X$ does not have the Liouville property. Then,  by homogeneity of the geodesic flow,  $\phi _1$ is not defined almost everywhere on $TX$ or it does not preserve the measure $\mathcal M$. Thus, we find a compact subset $K_1\subset X_{reg}$ and (by homogeneity) a Borel subset $A\subset T^1 K_1$ such that either $\phi_1$ is not defined on almost all of $A$ or $\mathcal M (A) > \mathcal M(\phi (A))$.  Replacing $A$
by a smaller  subset, if necessary,  we find in both cases some $\epsilon >0$, such that   $$\epsilon < \mathcal M (A) - \mathcal M(\phi_1 (A)) \; .$$
Since $$\phi_1 (A)=2\cdot \phi _2 \Big (\frac 1 2 \cdot A \Big)=2\cdot \Big(\phi _1 \circ \phi_1 \Big (\frac 1 2 A \Big ) \Big ) \; ,$$
we deduce
\begin{align*}
(\tfrac 1 2)^n \cdot \epsilon &\leq  \mathcal M \big(\tfrac 1 2\cdot  A \big) - \mathcal M \big( \phi_1 (\tfrac 1 2\cdot  A ) \big) +
\\
&\quad+\mathcal M \big(\phi_1 (\tfrac 1 2\cdot  A )\big)
 - \mathcal M\big (\phi_1 \big (\phi _1 (\tfrac 1 2\cdot  A )\big) \big).
\end{align*}
Thus,  taking  either $A_{\frac 1 2} := \frac 1 2  \cdot A$ or  $A_{\frac 1 2} := \phi_1 (\frac 1 2 \cdot A)$ we infer
$$\epsilon \cdot \Big(\frac 1 2 \Big )^{n+1}  < \mathcal M \big(A _{\frac 1 2} \big) - \mathcal M \big( \phi_1 (A _{\frac 1 2}) \big) \; .$$
The set $A_{\frac 1 2}$ constructed above is contained in $T^{\frac 1 2} K_{\frac 1 2}$, where $K _{\frac 1 2}  =B_{\frac 1 2} (K_1)$.
%is the closed tubular neighborhood  of
%radius $\frac 1 2$ around $K$.

Iterating the above procedure we obtain,  for $r_m =\frac 1 {2^m}$,   a subset $A_{r_m} \subset T^{r_m} K_m$  where  $K_m =B_{r_m} (K_{m-1})$
 %is the closed tubular neighborhood around $K_{m-1}$ of radius $r_m$
 and such that \eqref{eq:m} holds true.
The claim follows now since all $K_m$ are contained in the set $B_1 (K_1)$, whose closure is  compact, by completeness of $X$.
\end{proof}


\begin{rem}
The completeness of the space $X$ is used in the proof of \tref{thmmain} only once, namely in the last line of the above proof.
\end{rem}

\subsection{Relation with the mm-boundary}
Before proceeding to the proof of  \tref{thmmain}, we interpret the deviation measures $v_r$ from \eqref{eq:first} in suitable geometric terms.
Let $K\subset X$ be    measurable   and let $r>0$ be arbitrary.
Since $\mathcal H^n (X\setminus X_{reg} )=0$, we have
$$\mathcal M (T^{r} K) =\int _K \omega _n \cdot r^n  =\omega _n \cdot r^n \cdot \mathcal H^n (K) \; .$$
Denote now by $U^r(K)$
%\footnote{\color{red} Should we use $U^{<r}$? also, change $T^r$ to $T^{<r}$ everywhere?  \color{green} I have again decided against this correct but cumbersome notation, suggested by Tosha}
the set of all pairs $(x,y)\in X\times X$ with $x\in K$ and $d(x,y)<r$.
By Fubini's theorem the set $U^r(K)$ is $\mathcal H^n \otimes \mathcal H^n =\mathcal H^{2\cdot n}$ measurable and we have
$$\mathcal H^{2n} (U^r (K))= \int _K b_r (x) \cdot  d\mathcal H^n (x) \;. $$
Taking both equations  together, we see that the signed measure $v_r$ expresses the difference between $\mathcal H^{2n}$ and $\mathcal M$. More precisely,
\begin{equation} \label{eq:compare}
v_r (K) = \frac 1 {\omega_n \cdot r^n} \Big(\mathcal M (T^r K)- \mathcal H^{2\cdot n} (U^r (K)) \Big)\;.
\end{equation}



Now we can prove the following result,  reformulating \tref{thmmain}.
\begin{thm} \label{reform}
If an Alexandrov space $X$ has trivial mm-boundary then  it has the Liouville property.
\end{thm}

\begin{proof}
Assume that $X$ does not have the Liouville property and consider the compact subset $K \subset X$, the positive numbers $\epsilon, r_m$ and the Borel
subsets $A_m\subset T^{r_m} K$ provided by \lref{infini} above.

Let $Y$ be the closure of $B_1 (K)$.
Let again $D\subset TX$ be the set of all vectors at which the exponential map is defined.
  For $r>0$, denote  by $N^r$ the intersection of $D$ with $T^r Y$ and consider the "total exponential map"
$E\:N^r \to X\times X$ given by $$E(v)= (\pi (v), \pi (\exp (v))) \; .$$
 As above,  let $U^r =U^r(Y)$ be  the set of all
  pairs $(y,x) \in X\times X$ with $y\in Y$ and $d(x,y)<r$. Then
\begin{equation} \label{eq:image}
E(N^r) =U^r \; .
\end{equation}
Moreover, for any  fixed $x \in Y$, the restriction of $E$ to $D_x \cap N^r$ is
a $(1+ C r^2)$-Lipschitz continuous map from $D_x\subset T_xX$ onto the set  $U^r \cap (\{x \} \times X)$.
Thus, for all sufficiently small $r$, and any Borel subset $S\subset D_x \cap N^r$, we have
$$\mathcal H^n (E(S)) \leq (1+4n C r^2)\cdot  \mathcal H^n (S)\; .$$
 Using the definition of the measure $\mathcal M$ and  Fubini's formula for the product measure
$\mathcal H^{2\cdot n} =\mathcal H^n \otimes \mathcal H^n $ on $X\times X$ we obtain for any $\mathcal M$-measurable subset $S$ of $N^r$
\begin{equation} \label{eq:contract}
\mathcal H^{2n} (E(S)) \leq (1+4nC r^2) \cdot \mathcal M(S)\; .
\end{equation}
Due to \eqref{eq:compare}, the triviality of the mm-boundary of $X$ implies
$$\lim _{r\to 0} \frac 1 {r^{n+1} } |\mathcal M(T^{<r} Y) -\mathcal H^{2n} (U^r)| =0 \;.$$
 Thus, up to terms of higher order, the  map $E$ does not increase the measure of subsets, but the total mass of the image
coincide with the total mass of the target.
Thеrefore, $E$ is measure preserving up to terms of higher order.
% \footnote{Is this explanation helpful? \color{red} too long IMO. I shortened it a bit, Vitya}
More precisely,   combining the last two inequalities we obtain for every  $\delta >0$ the existence of some $s>0$ with the following property. For all $0<r<s$ and all measurable subsets $S\subset N^r$
	we have
	\begin{equation}  \label{eq:almostall}
	\mathcal M (T^r Y)- \mathcal M (N^r) < \delta r^{n+1} \;   \text{and}  \;   |\mathcal H^{2n} (E(S)) - \mathcal M  (S)|  < \delta r^{n+1} \;.
	\end{equation}
%	and  for all measurable subsets $S\subset N^r$ we have
%	\begin{equation}  \label{eq:almostiso}
%	|\mathcal H^{2n} (E(S)) - \mathcal M  (S)|  < \delta r^{n+1} \;.
%\end{equation}
For any measurable subset $S\subset N^r \cap TK$  we now claim
\begin{equation} \label{eq:finally}
|\mathcal H^{2n} (E(S)) - \mathcal M (\phi _1 (S))|  <  2 \delta r^{n+1} \; .
\end{equation}
In order to prove \eqref{eq:finally}, let $S^+$ be the subset of all vectors $v\in S $ for which $\phi_1 (v)$ exists  and is contained in $TX_{reg}$.
For all $v\in S^+$, we have $-\phi _1 (v) \in N^r$ and,   due to \eqref{eq:symm},
$$E(-\phi_1 (v)) =J(E(v)) \;.$$
The involution  $I(v)=-v$ is $\mathcal M$-preserving on $TX_{reg}$.  And the involution
 $J\:X\times X\to X\times X$ given by $J(x,y)=(y,x)$ preserves $\mathcal H^{2n}$. Therefore, from \eqref{eq:almostall} we deduce
 \begin{equation} \label{eq:final+}
 |\mathcal H^{2n} (E(S^+)) - \mathcal M (\phi _1 (S^+))|  <   \delta r^{n+1} \;.
\end{equation}
On the other hand, by construction,
 $$ \mathcal M (S\setminus S^+ )=0 \; \text{and} \; \phi _1 (S\setminus S^+)\cap TX_{reg} =\emptyset \; .$$
  Hence, applying \eqref{eq:almostall}, we see
 $$|\mathcal H^{2n} (E(S))- \mathcal H^{2n} (E(S^+))|< \delta r^{n+1} \; \text{and} \; \mathcal M (\phi _1 (S\setminus S^+)) =0 \;.$$
 Together with \eqref{eq:final+}  this finishes the proof of \eqref{eq:finally}.


Coming back to our subsets $A_m\subset T^{<r_m} K$ we have
 $$\epsilon \cdot r_m^ {n+1} \leq \mathcal M ( A_m) - \mathcal M( \phi _1 (A_m)) \leq \mathcal M ( A_m) - \mathcal M( \phi _1 (A_m\cap N^r)) \,.$$
Setting  $S_m=A_m\cap N^{r_m}$ we estimate the right hand side from above by
$$|\mathcal M (A_m) -  \mathcal M(S_m)|
+|\mathcal M (S_m )  - \mathcal H^{2n} (E(S_m) )| + | \mathcal H ^{2n} (E(S_m)) - \mathcal M (\phi _1(S_m)|  \;.$$
Applying  \eqref{eq:almostall} and \eqref{eq:finally}  this expression is bounded above by $4 \delta r _m^{n+1}$, for all $m$ large enough. Taking the inequalities together we obtain,  $\epsilon \cdot r_m^ {n+1}< 4 \delta r _m^{n+1}$ for all $m$ large enough.
Since $\delta$ is arbitrary small, this provides a  contradiction  and  finishes the proof of \tref{reform}.
% and therefore of \tref{thmmain}.
\end{proof}









%\subsection{Another measure}

%We will need a simple lemma:

%\begin{lem} \label{anothermeasure}
%Let $L,n$ be given. There is some $\bar L= \bar L (L,n)$ with the following properties.
%Assume that $Lr^n \geq \vol (B_r (x)) \geq  1/L r^n$ for all $x \in X$ and all small
%$4r\leq r_0$.  Let $K$ be a positive measure on $X$. Let $A\subset X$ be given, with finite $K (B_{2r} (A))$.
%Then $\int _A K(B_r (x)) d\vol (x) \leq \bar L r^n \cdot K (B_{2r}(A))$.
%\end{lem}

%\begin{proof}
%Cover $A$ by balls $B_i$ of radius $2r$, such that the  balls with the same centers and radii $r$  still cover $A$ and such that
%the intersection multiplicity of the covering $B_i$ is bounded by some
%$L'=L' (L)$. Then estimate  all $K(B_r (x))$ by $K(B_i)$ for some  for some $i$,
%with $x\in B_i$. We deduce

%$$\int _A K(B_r (x)) d\vol (x) \leq \int _A  K(B_i) \cdot Char _{B_i} (x)  d\vol (x) \leq $$

%$$\leq \Sigma  K(B_i) L (2r)^n \leq 2^n L L' K(B_{2r} (A))$$
%\end{proof}


%\subsection{Regular-Singular Decomposition}

%\begin{prop} \label{regsing}
%Let $X$ be a space with another positive measure $K$ (absolute value of scalar curvature). Assume that there are constants
%$C$ (measure of singular locus), $L$ (some growth constant), $l$ (equal to 1 or 2 for us), $\delta ,r_0$
%(some thresholds) such that the following holds true.  For any $r<r_0$ there is a decomposition
%$X=X_r \cup Y_r$ with $\vol (Y_r) \leq C\cdot r^l$. For any $x\in X_r$ and any  $s\leq r$ the volume of
%$ (B_s (x)) $ is bounded by $L^{\pm 1} s^n$ and for any $x\in X$ it is bounded from above by $L\cdot s^n$.
%  Assume further, that for any  $x\in X_r$ the inequality
%$K(B_{Lr} (x)) \leq \delta  r^{n-l}$ implies $|\vol (B_r (x) ) - b_n r^n| \leq L \cdot K ( B_{Lr} (x)) \cdot r^l$.

%Then there is some number $\bar L$, such that for all small $r$ and all $A\subset X$ we have
%$\|v_r (A)\| \leq \bar L \cdot (K(B_{2Lr} (A)) +C) \cdot r^l$.
%\end{prop}




%\begin{proof}
% Set  $S^r := \{ x\in A \cap X_r| K(B_{Lr} (x)) > \delta  r^{n-l} \}$.
% Let $N_1$ be the maximal $2Lr$-net in $S^r$. Then $|N_1| \leq K (B_{Lr} (A)) /(\delta _0  r^{n-l})$ elements.
% Thus $\vol (S^r)$ is bounded from above by $2^n \cdot L ^{n+1} \cdot \delta _0 ^{-1} \cdot r^l \cdot K(B_{Lr} (A)) $.
% Therefore $|v_r |(S^r) \leq \bar L_1 \cdot K (B_{Lr } (A)) \cdot r^l$.

% On the other hand, $\vol (A \cap Y_r) \leq C \cdot r^l$.  Thus
% $|v_r| (A\cap Y_r) \leq \bar L_2 \cdot C \cdot r^l$.

% Finally, for any $x$ in $A^+ := A\setminus (S^r \cup  Y^r)$, we have
% $|\vol (B_r (x))- b_n r^n | \leq L \cdot K(B_{Lr} (x)) \cdot r^l$.

% Thus $r^n \cdot |v_r| (A^+) \leq \int _{A^+} L \cdot K(B_{Lr} (x)) \cdot r^l d\vol (x)
% \leq \bar L_3 \cdot r^{n+l} \cdot K (B_{2Lr} (A^+))$, by \lref{anothermeasure}.

% Summing up, we obtain the desired inequality.
%  \end{proof}

%\subsection{Limiting case}
%Under the assumptions of the previous section, let $\delta$  be fixed. Assume that as $r$ goes to $0$, all subsets $X_r$ are open (this assumption can be easiely achieved by enlarging $X_r$ slightly). Assume further that $X_r \subset X_s$ for
%all $s\leq r$. Set $X_{reg} = \cup X_r$ and $Y_0 = \cap Y_r$.  Set, finally,
%$C''=\limsup _{r\to 0}  \frac {vol (Y_r)} {r^l}$.

% From \pref{regsing}, we conclude that the measures $\frac {v_r } {r^l}$ are
%uniformly bounded.  Let $\mu$ be any weak limit
%$\mu = \lim _{r_i \to 0} (v_{r_i}  / r_i ^l)$.  Then for some constant
%$L$ depending on the privious $L$, the measure $\mu$ satisfies the following:

%1) $\mu (Y_0) \leq L C'$;

%2) $\mu$ is bounded on $X_{reg}$ by $L \cdot |K|$.


\subsection{Quasi-geodesics} \label{subsec:quasi}
Finally, we discuss some relations with quasi-geodesics, referring the reader to \cite{Petsemi} for the basic properties of such curves.
Recall, that whenever a unit speed minimizing   geodesic $\gamma _v :[0,a] \to X$   start at a point $x$ in the direction $v$ then this is the unique quasi-geodesic defined on the interval $[0,a]$, \cite{PP}, p.8, thus the same statement is also true for  (local) geodesics $\gamma _v$.


Using this and the fact that a limit of quasi-geodesics is a quasi-geodesic, it is not difficult to conclude that the partial geodesic flow $\phi:\mathcal F\cap TX_{reg} \to TX_{reg}$ defined above is continuous, slightly strengthening Subsection \ref{subsec:measur}.


As in Subsection \ref{subsec:tb}, we have a canonical measure $\mathcal M _1$   on the unit tangent bundle $\Sigma X \subset TX $ of $X$, which we also call the Liouville measure. Whenever
$X$ has the Liouville property, then the geodesic flow is defined $\mathcal M_1\otimes \mathcal H^1$-almost everywhere on $\Sigma X \times \R$ and preserves $\mathcal M_1$.  In this case    for $\mathcal M_1$-almost each unit direction there exists exactly one quasi-geodesic starting in this direction.

Let now $X$ be an  Alexandrov space with topological boundary $\partial X$ and let $Z$ be the doubling $X\cup _{\partial X} X$, which
is an Alexandrov space without boundary, \cite{P2}. Quasi-geodesics in $X$ are exactly the projections of the quasi-geodesics in $Z$ under
the canonical folding $f:Z\to X$.   From this we deduce that if $Z$ has the Liouville property, then $\mathcal M_1$-almost each  direction $v\in \Sigma X$ is the starting direction of a unique infinite quasi-geodesic in $X$. Moreover, in this case,  the corresponding quasi-geodesic flow preserves $\mathcal M_1$.

Finally, as an application of \tref{thmmain} and \tref{alexandrovthm} we see that the above assumptions are fulfilled whenever the complement $X\setminus\partial X$ has trivial mm-boundary. Indeed, in this case the mm-boundary of $Z$ must be concentrated on $\partial X \subset Z$, hence it must be trivial by \tref{alexandrovthm},(3).



\section{Surfaces with bounded integral curvature in the sense of Alexandrov} \label{sec:surface}
\subsection{Preparations} We assume here some  familiarity  with the theory of surfaces with bounded integral curvature, cf. \cite{AZ} and
 \cite{Reshetnyak-GeomIV}.  Such a surface $X$ is a locally geodesic metric space, homeomorphic to a two-dimensional surface.  It has Hausdorff dimension $2$  and  the Hausdorff measure $\mathcal H^2$ is a Radon measure  on $X$.   There is another canonical signed Radon measure on $X$, the so called curvature measure $\Omega$, \cite[Section 8]{Reshetnyak-GeomIV}. Unlike when talking about Alexandrov spaces \emph{we will not assume that $X$ is complete}.
 We will derive  \tref{intsurface} as a consequence of the following local and non-exact version of a theorem of M. Bonk and U.Lang, \cite{Bonk-Lang}, relating the curvature measure $\Omega$ to the volume of balls.

\begin{lem} \label{lem:bl}
There exists some $\delta _0>0$ with the following property.
Let $X$ be a surface with bounded integral curvature and let  $\Omega\in \mathcal M(X)$ be its   curvature measure. Let $x\in X$ be a point and let $r>0$ be such that
$\bar B_{r} (x)$ is compact,  $| \Omega| ( B_{r} (x) )  < \delta _0$ and $\bar B_{r} (x)\subset U$ where $U$ is an open subset of $X$ homeomorphic to $\R^2$.
Then $$|1- \frac {b_r(x)}  {\pi r^2} | \leq 3 \cdot |\Omega |( B_{r} (x)) \; .$$
\end{lem}


\begin{proof}
Set $\delta = |\Omega |( B_{r} (x))$. By continuity, it is sufficient to prove that $|1- \frac {b_s(x)}  {\pi s^2}| \leq 3 \cdot \delta $ for any $s<r$.  Using approximations of
the metric on $U$ by polyhedral metrics \cite[Theorem 8.4.3, Theorem 8.1.9]{Reshetnyak-GeomIV},  it is sufficient to prove this  statement for polyhedral metrics. Thus we assume from now on that $X$ is polyhedral  and homeomorphic to $\R^2$.

\emph{Claim:} There exists a complete polyhedral surface $\hat X$ homeomorphic to a plane,  which contains a copy of $B_s (x)$ and such that
the curvature measure $\hat \Omega$ of $\hat X$  satisfies $|\hat \Omega | (\hat X) < 3\delta$.  Once the claim is proven  and $3\delta < 2\pi$,
\cite{Bonk-Lang} provides us an $L$-biLipschitz map $f:\hat X\to \R^2$ with $L\leq 1+\frac {3\delta} {2\pi -3\delta}$.
For sufficiently small $\delta$, we would only need to apply \eqref{eq:bilip} to finish the proof of the lemma.

It remains to prove the existence of $\hat X$. We take some $r>t>s$ and consider the compact  metric ball $\bar B_t (x) \subset B_r (x)$. We may assume that  the boundary $S_t=S_t (x)$ of
$\bar B_t (x)$ does not contain singular points of $X$. Due to \cite[Theorem 9.1, Theorem 9.3] {Reshetnyak-GeomIV}, the boundary $S_t$ is a (piecewise smooth) Jordan curve, once $\delta _0< 2\pi$, and the negative part $\kappa ^-$ of the geodesic curvature $\kappa$ of $S_t$  satisfies $|\kappa ^-| (S_t)  \leq \delta$.  Since $U$ is homeomorphic to a plane this implies that $\bar B_t (x)$ is homeomorphic to a closed disk $\bar D^2$ in $\R^2$.


We find a polygonal Jordan curve $\Gamma$ in $B_t(x)$  approximating $S_t$  such that the negative part of the geodesic curvature of $\Gamma$ is smaller than $2\delta$.  Consider the closed Jordan domain  $Y$ bounded by $\Gamma$, which can be assumed to contain $B_s(x)$.
Now we glue to $Y$ along any edge of $\Gamma$ a flat half-strip.  The boundary of the arising polyhedral surface  consists of pairs of rays
$\gamma _i ^{\pm}$  emanating from the vertices $V_1,..,V_k$ of $\Gamma$. The rays $\gamma _i^{\pm}$ enclose an angle equal to $2\pi- \alpha _i$, where $\pi-\alpha _i$ is the angle of $\Gamma$ at $V_i$ measured in $Y$.   In order to finish the construction of $\hat X$ we glue  a flat sector of
angle $\alpha _i$ between $\gamma _i ^{\pm}$, if $\alpha _i >0$ and we glue $\gamma _i^{\pm}$ together if $\alpha _i \leq 0$.
Since $Y$ was a polyhedral disc, the arising space $\hat X$ is a complete polyhedral plane.
 All of the singularities of $\hat X$  are contained in $B_s(x) \cup \{ V_1,...., V_k\}$. Moreover, by construction, the curvature measure
   $\hat \Omega$ of $\hat X$ satisfies
   $$\hat \Omega  (V_i) = \min \{ 0, \alpha _i \} \;.$$
  We deduce,
  $$|\hat \Omega | ( \hat X) = |\Omega (B_s (x))| + |\hat \Omega | ( \Gamma ) \leq \delta  + |\kappa ^-| (\Gamma ) < 3 \delta \;.$$
    This finishes the proof of the claim and of \lref{lem:bl}.
\end{proof}



 \subsection{Local finiteness of mm-curvature}
 Now we are in position to prove
 the following generalization of  \tref{intsurface}.
 \begin{thm}\label{intsurface1}
 Let $X$ be an Alexandrov surface with integral curvature bounds.  Then,  equipped with the Hausdorff measure $\mathcal H^2$, the space $X$ has locally finite mm-curvature.
  \end{thm}
 %
 \begin{proof}
 Let again $\Omega$ denote the curvature measure of $X$. Let $\delta _0>0$ be sufficiently small and satisfy the conclusion of  \lref{lem:bl}.
 The statement  of \tref{intsurface1} is local, so we need to prove it only in a small neighborhood of any point.
 Thus we may and will assume that there is a point $x_0\in X$
 such that $| \Omega| (X \setminus  \{ x_0 \} )< \delta _0$ and that $X$ is homeomorphic to a plane.



  Let  $A\subset X$  be a compact subset. Choose some $\epsilon >0$ such that the closure of  $B_{2\epsilon} (A)$ in $X$ is compact and such that,
  for any $0<2r<\epsilon$  the inequality  $\mathcal H^2 (B_{2r}(x_0)) < \frac 1 {\epsilon} r^2$ holds true,   cf. \cite[Lemma 8.1.1]{Reshetnyak-GeomIV}.

 Let  $r<\epsilon$    be arbitrary.
 For any $x\in B_{2r} (x_0)$ we have $$b_r (x)  =\mathcal H^2 (B_r (x)) \leq \mathcal H^2 (B_{3r} (x_0)) \leq \frac 1 {\epsilon} r^2 \; .$$
For any $x\notin B_{r} (x_0)$ we have $| \Omega |(B_r (x)) < \delta _0$. Thus, by Lemma~\ref{lem:bl}, $$|1-\frac {b_r(x)}  {\pi r^2} | \leq 3 |\Omega | ( B_{r} (x))\; .$$
For the  deviation measures $v_r$ from \eqref{eq:first} we estimate:
 $$|v_r| (A\cap B_{2r} (x_0))) \leq |v_r| (B_{2r} (x_0))  \leq (1+ \frac 1 {\epsilon}) \cdot \mathcal H^2 (B_{2r} (x_0)) \leq
 (1+ \frac 1 {\epsilon}) \cdot \frac 1 {\epsilon} r^2 \; .$$
On the other hand,
 %for any $x\in A\setminus B_{r} (x_0)$ we have $\bar \Omega (B_r (x)) < \delta _0$, hence $|\frac {b_r(x)}  {\pi r^2}  -1| \leq 5 |\Omega | (\bar B_{r} (x))$.  Therefore,
$$|v_r| (A\setminus B_{2r} (x_0)) \leq \int _{A \setminus B_{2r} (x_0) } 3 \cdot |\Omega| (B_r (x)) \cdot  d\mathcal H^2 (x)  \leq $$
$$\leq 3 \int _{B_r (A\setminus B_{2r} (x_0))} \mathcal H^2 (B_r(x))  \cdot d|\Omega | (x) \; ,$$
%where $Q= B_r (A\setminus B_{2r} (x_0)$, and
where we have used \lref{lem:exchange} in the last step.
 For any $x$  contained in the domain of integration of the last integral, we have $\mathcal H^2 (B_r(x))  =b_r (x) \leq 2\pi  r^2$, by \lref{lem:bl}, once $\delta _0$ has been chosen to be sufficiently  small.
We deduce $|v_r| (A\setminus B_{2r} (x_0)) \leq 6 \pi \delta _0 \cdot r^2$.

Thus, for some constant $C=C(\epsilon)$ and all $r<\epsilon$, we obtain
$$|v_r|(A) \leq C\cdot r^2 \; .$$
This finishes the proof of the theorem.
\end{proof}









\section{Convex hypersurface} \label{sec:hyper}
In this section we are going to prove \tref{thmfirst}. The proof will follow from \tref{thmmain} by comparing the mm-boundary with the mean curvature measure on convex hypersurfaces.
It is possible  to deduce the theorem without a reference to \tref{alexandrovthm}, from \lref{lem:mean} alone, but using  our main technical result \tref{alexandrovthm}
shortens the proof.   All results in this section are local. For simplicity, we nevertheless assume all convex hypersurfaces to be closed in $\R^{n+1}$. They will always be equipped with the induced intrinsic metric, not with the restricted distance function from $\mathbb R^{n+1}$.
We assume in here some familiarity with the  basics of the theory of convex functions and convex geometry.
\subsection{Mean curvature}
Let $X^n$ be a  convex hypersurface in $\R ^{n+1}$.  Recall that there exists a Radon measure $K$ on $X$, called the mean curvature measure, which has the following properties, cf. \cite{Schneider}, \cite{Fedcurvature}.
For smooth hypersurfaces $X$,  we have $K=\kappa \cdot \mathcal H^n$, where $\kappa$ is the  usual mean curvature function of $X$.
   The mean curvature measure is stable under Hausdorff convergence of convex hypersurfaces in $\R^{n+1}$. If the hypersurface is rescaled by $\lambda$, the mean curvature $K$ is rescaled by $\lambda ^{n-1}$.

We call a point $x$ in the convex hypersurface $X$ a smooth point if there is a unique supporting hyperplane of $X$ at this point.
For  any smooth point $x\in X$, any sequence $x_j\in X$  converging to $x$ and any sequence of positive numbers $t_j$ converging to $0$, the
sequence of convex hypersurfaces  $X_j$ obtained from $X$ by the  dilatation by the factor $\frac 1 {t_j}$ centered at the point $x_j$ converges
to the tangent hyperplane of $X$ at $x$. Using   stability of the mean curvature measures  $K$, vanishing of $K$ on   flat hyperplanes and
 the behavior of $K$ under rescalings gives us the following conclusion.
\begin{lem} \label{lem:compsm}
Let $X$ be a convex hypersurface in $\R^{n+1}$. Let $A$ be a compact set of smooth points in $X$ and let $\delta >0$ be arbitrary.
Then there exists some $t>0$ such that for any  $y\in B_{t} (A)$ and any $0<r<t$ we have $K(B_r(y))  \leq \delta \cdot r^{n-1}$.
\end{lem}
 Thus, the following lemma applies to all small balls in a neighborhood of any smooth point.






%The following lemma is basic:

\begin{lem}  \label{lem:mean}
There exist numbers $\delta_0,C>0$ depending only on $n$ with the following property. Let $X$ be a convex hypersurface in $\R^{n+1}$.
Let $x\in X$ be a point and $r>0$ be such that the mean curvature $K$ satisfies  $K(B_{6r} (x))  < \delta \cdot r^{n-1}$ with $\delta <\delta _0$.
Then
\begin{equation} \label{eq:lemma}
| 1 -  \frac {b_r (x)} { \omega _n r^n}| < C\cdot \delta \cdot K( B_{6r} (x)) \cdot r^{1-n} \; .
\end{equation}
\end{lem}


\begin{proof}
By rescaling, it suffices to prove the existence of $\delta _0, C>0$ such that the lemma holds true for $r=1$.
By approximation, it is enough to prove the result for smooth convex hypersurfaces.


We fix a sufficiently small $\epsilon _0>0$.   The mean curvature vanishes on $B_6 (x)$ if an only if $B_6(x)$ is contained in a flat hyperplane.
 Due to  the stability of $K$ under convergence,      the ball $U=B_{5} (x) \subset X$ must be  arbitrary close to a flat hyperplane in $\R^{n+1}$,
once $\delta _0$ has been chosen small enough.  Thus, we may assume
 that  the tangent hyperplanes  to points in $U$  are $\epsilon_0$-close to  the tangent
space  $W =T_xX \subset \R^{n+1}$.  Therefore, $U$ is a graph  $U=\{(x,f(x))\}$ of a convex function $f:V\to \R$ defined on an open subset $V\subset W$. Moreover, $V$ contains the ball of radius $4$ in $W$ around $x$. Let $O$
be the ball of radius $2$ in $W$ around $x$. Set
$$a :=\sup \{ |\nabla f (y)| \; ; y\in O \} \; .$$
If $\delta _0$ is small enough, then $a<\epsilon_0$.  The orthogonal projection $P:U\to V$ is $1$-Lipschitz and the restriction of the inverse $P^{-1}$ to $O$  has Lipschitz constant  $$\sqrt {1+ a^2}  \leq 1+a^2 \leq 1+\epsilon_0 ^2 \; .$$
Applying \eqref{eq:bilip} we only need to prove that $a < C \cdot \delta $ for a constant $C$.

%In the following we denote by $C_1,C_2...$ positive constants depending only on $n$.
Denote by $|D^2f|$ the largest eigenvalue of the Hessian $D^2f$.  Since $f$ is convex and $\epsilon _0$ small enough,  the mean curvature $\mathcal \kappa (x)$ at the point $(x,f(x))$ of the graph $U$ of $f$ satisfies $\kappa (x) \geq \frac n 2  \cdot  |D^2 f|$.  Hence, the conclusion  follows from the following statement.


\emph{Claim:} Let $f:B \to \R$ be a smooth convex function on the open ball $B=B_4 (0)\subset \R^n$.
If $f(0)=|\nabla f(0)| =0$ then, for some $C=C(n)>0$,
$$\sup_{y\in B_2 (0)} |\nabla f (y)| \leq C\cdot \int _{B} |D^2f| \;.$$
By convexity, it is enough to find some $C=C(n)>0$ with
%The function $f$ is non-negative.
%Since the gradient of $f$ is controlled by the supremum of $f$ on a slightly larger ball, cf. \cite{Evans}, Theorem 6.7,
%it suffices to prove that
\begin{equation}  \label{eq:vitya}
\sup _{y\in B_3 (0) }| f(y)| \leq C \cdot \int _{B} |D^2f|\; ,
 \end{equation}
 cf. \cite[Theorem 6.7]{Evans}.

In order to verify \eqref{eq:vitya}, we can multiply the function $f$ by a constant and assume that
 $f$ takes its maximum on the closed ball $\bar B_3 (0)$ at the point $y_0$ and $f(y_0)=1$.  Convexity of $f$ implies that $|y_0|=3$.
 Since $f(0)=0$ and  $f$ is convex, we must have $f(y)\leq \frac 1 3$ for all $y\in B_1(0)$.
  %must be bounded by $\frac 1 3$,  $|\nabla  f | <\frac 1 2$ on the ball $B_1 (0)$, cf. \cite[Theorem 6.7]{Evans}.
 %Hence $f$ is bounded on the ball   $B_{\frac 1 2} (0)$ by $\frac 1 4$.

 By convexity and the choice of $y_0$, the restriction of $f$ to the supporting  hyperplane $H$ of $\bar B_3(0)$ at $y_0$ is bounded from below by $1$. Consider the ball  $S$ of radius $\frac 1 2$ in $H$ around $y_0$. For any point $z\in S$ consider the restriction
 $$f_z  (t)= f(z- \frac t 3 y_0 ) \;, t\in [0,6]$$  to the segment of length $6$   starting at $z$ orthogonal to $H$.
 Then $$f_z(0)\geq 1 \; ; \; f_z (3)\leq \frac 1 3 \; ; \; f_z (6) \geq 0 \;.$$
 Thus for some $t\in (0,3)$ we have $f_z'(t) \leq  -\frac  2 9$ and for some $t\in (3,6)$ we have $f_z'(t) \geq  - \frac  1 9$.
 Therefore
 $$\int _0 ^6 f_z'' (t)\, dt  \geq \frac 1 9 \; .$$
 Integrating over $S$ we obtain by Fubini's  theorem a uniform positive lower bound on $\int _{B} |D^2f|$.
 This finishes the proof of \eqref{eq:vitya}, of the claim and of Lemma.
\end{proof}




\subsection{The proof} Now we can provide the proof of
\begin{thm} \label{thmconv}
Let $X=X^n$ be a convex hypersurface in $\R^{n+1}$.  Then it has trivial mm-boundary.

\end{thm}
Together with \tref{intsurface} this finishes the proof of \tref{hypersurface}. In combination with \tref{thmmain} it also finishes the proof of \tref{thmfirst}.
% it is sufficient to show that $X$ has trivial mm-boundary. This claim is local on $X$.
%Taking a small neighborhood of an arbitrary point in $X$ we can extend it to compact convex hypersurface. Thus we may and will assume that $X$
%is compact.
\begin{proof}
%For any $r>0$ the signed measure $v_r$ on $X$ defined by \eqref{eq:first} is non-negative by the theorem of Bishop--Gromov.
 Since $X$ has locally finite mm-boundary by
\tref{alexandrovthm}, it suffices to prove that  any limit measure $\nu$ of a sequence $\frac 1 {r_j} v_{r_j}$ for $r_j\to 0$ must be the zero measure.
Let us fix such a limit measure $\nu$.  Due to \tref{alexandrovthm}, $\nu (A)=0$ for any Borel subset $A\subset X$ with $\mathcal H^{n-1} (A)<\infty$.
Let $Y\subset X$ be the set of smooth points of $X$.
%i.e. of points with exactly one supporting hyperplane.
The complement $X\setminus Y$ is a countable union of subsets with finite $(n-1)$-dimensional Hausdorff measure, \cite[Theorem 1.4]{Schneider}, therefore $\nu (X\setminus Y) =0$.   Therefore, it is sufficient to prove $\nu (A)=0$
for any compact subset $A\subset Y$.



Let $A$ be such a compact subset and  let $\delta>0$ be an arbitrary sufficiently small number. Consider a positive $1>t>0$ provided by \lref{lem:compsm}.  Let $U$ be the open set $B_t (A)$. Making $t$ smaller, if needed, we may assume that $K(B_{7t}(A))$ is bounded by a
 fixed constant $C_1$ independent of $\delta$.

 Let $0<r<t$ be arbitrary. Then, by \lref{lem:mean},  for any $x\in U$ we have \eqref{eq:lemma}.
Therefore, $$|v_r| (U)\leq \int _U   C\cdot \delta \cdot r^{1-n}\cdot  K (B_{6r} (y)) \cdot d\mathcal H^n (y) \leq  $$
 $$\leq C\cdot \delta \cdot  r^{1-n} \cdot \int _{B_{7t} (A)} \mathcal H^n (B_{6r} (y)) \cdot dK(y)  \leq C\cdot \delta \cdot  r^{1-n} \cdot (6r) ^n \cdot K(B_{7t} (A)) \;,  $$
where we have used \lref{lem:exchange} in the second and the theorem of Bishop--Gromov in the last inequality.  Hence
$$|\nu| (A) \leq |\nu| (U) \leq C\cdot \delta \cdot 6^n \cdot K(B_7 (A)) \;. $$
Since $\delta$ can be chosen arbitrary small, we obtain $|\nu| (A)=0$.

This finishes the proof of the claim and, therefore, of \tref{thmconv}.
\end{proof}








%\subsection{Generalization}
%The following more general statement seems to follow in the same way. Maybe it should be better steted after the section about
%BV-metrics.

%\begin{defn}
%A $DC$-submanifold $N$ of a smooth submanifold $M$ is a subset that can be locally around each point represented as a graph
%$N=(x,f(x))$ for a $DC$-map $f\:\R ^m \to \R ^n$.
%\end{defn}

%Main examples are $C^{1,1}$ submanifolds and convex hypersurfaces in Riemannian manifolds.
%Using the theorem of Nash, one can (isometrically) consider any $DC$-submanifold of a Riemannian manifold
%as a $DC$-submanifold of a euclidean space.

% It seems to me that everything explained above workes well in this more general situation and one obtains:

% \begin{prop}
% Let $N$ be a $DC$-submanifold of a smooth Riemannian manifold. Then $N$ has trivial mm-boundary.
% \end{prop}

\section{An integral inequality for Riemannian metrics}\label{sec-BV-estimate}
\subsection{The smooth case} We start by estimating from above the deviation measure $v_r$  on a smooth Riemannian manifold in terms of the first
derivatives of the metric.  We do not know how to prove a similar estimate from below, see  Problem \ref{qe:BV}. However, for the applications
to Alexandrov spaces discussed in the next section, the  estimate from below is a consequence of the theorem of Bishop--Gromov.

For a smooth Riemannian metric $g$ defined on an open subset $U\subset \R^n$ we denote by $|g'|:U\to [0,\infty)$ the sum $\Sigma_{i,j,k}|\frac \partial {\partial \,x_k} g_{ij}|$.

%of the absolute values of the first partial  derivatives $|\frac \partial {\partial \,x_k} g_{ij}|$ of the coordinates  of $g$.

\begin{prop}\label{prop-smooth}
There exists a constant $C=C(n)>1$ with the following property.
Let $U\subset \R^n$ be an open subset with a smooth Riemannian
metric $g$ which is $(1+\frac 1 C)$-biLipschitz to  the background Euclidean metric.
%t $g_0$ denote the flat background metric and assume $\|g_x-g_0\| \leq \epsilon$ for some fixed small
%$\epsilon$.
  Let $A\subset U$ be a Borel subset. Let $r>0$ be such that $B_{2r} (A)$ is relatively compact in $U$.
Then  $$v_r (A) \leq C \cdot r \cdot \int _{B_{2r} ( A)} |g' | \;.$$
\end{prop}



 \begin{proof}
Throughout the proof we will denote by $C_i$ various (explicit) constants which depend only on $n$.
%Let us first set $C_0=100n$ and let $\eps=\frac 1 C_0$.  To simplify notations throughout this proof we will denote $B_{2r} (A)$ by $A_r$.

We will use the following notations. By $|\cdot|$ and $\mathcal L^n$ we denote respectively the norm and the Lebesgue measure on the background $\R^n$.
 For $x\in U$ we denote by $g_x$ the Riemannian tensor at the point $x$ and
 by $|\cdot |_x$ the corresponding norm. % $\R^n$ will be considered with the Euclidean metric coming from $g_0$
% and the corresponding Lebesgue measure $\mathcal L^n$.
The Hausdorff measure of the Riemannian metric
 $g$ has the form $u\cdot \mathcal L^n$, with $u=\sqrt {det (g_{ij})}$.




For $x\in \R^n$, we consider the function $K :U\to [0,\infty)$ given by
$$K (x)=\sup _{|v|_x =1}\,  \frac d {dt}\Big |_{t=0} |v| _{x+tv} \; .$$
%Once $C$ has been chosen small enough we have $\frac 1 2 < u <2$ on $U$.
  By smoothness of the determinant and the square root,
we find a constant $C_1$ such that  for all $x\in U$  we have
\begin{equation} \label{eq:Kg}
|u' (x)|  \leq C_1\cdot  |g' (x)|  \; \;  \text{and}  \; \;  K (x)  \leq C_1 \cdot  |g' (x)|  \;.
\end{equation}
 We fix $A\subset X$ and $r>0$ as in the formulation of the proposition. For $x\in U$ denote by $B_x$
 the metric ball $B_r(x)$ in $U$.  By $B^x$ we denote the metric ball of radius $r$ in the Euclidean  norm $|\cdot |_x$.  In this Euclidean metric the ball $B^x$ has measure
 $$\omega _n r^n =u(x)\cdot \int _{B^x} d\mathcal L^n \;.$$
Thus, in order to estimate the deviation measure $v_r$, we only need to control the summands on the right of the following  inequality:
\begin{equation} \label{eq:summand}
\omega _n \cdot r^n - b_r (x) \leq  u(x) \cdot \mathcal L^n(  B^x\setminus B_x)   + \int _{B_x}  |u(x)-  u(y)| \cdot
d\mathcal  L^n (y)   \; .
\end{equation}



We may assume that the biLipschitz constant $1+\frac 1 C$ is close enough to 1,  so that $\frac 1 2 < u <2$. Moreover, we may assume $B_x$ and $B^x$ are contained in the ball of radius $\frac 4 3 r$ around $x$ with respect to the background Euclidean metric.


In order to bound the first summand,  for $x\in A$ and $|v|_x=1$,  we set $l^x_v$ to be the length of the segment $[x,x+v]$
in the Riemannian metric $g$.  Then we compute
$$l_v ^x -r= \int _0 ^r |v|_{x+tv} \, dt - \int _0 ^r |v|_x \, dt   \leq \int _0 ^r (\int _0 ^t  K(x+sv) \, ds)\, dt \leq $$
$$\leq \int _0 ^r (\int _0 ^r   K(x+sv) \, ds)\, dt = r\cdot \int _0 ^r   K(x+sv) \, ds  \; .$$
Observe now that the intersection of $B^x\setminus B_x$ with the ray starting in $x$ in the direction of $v$ has $\mathcal H^1$-measure
( with respect to the norm $|\cdot |_x$)
at most $2\cdot (l_v^x -r)$, once the biLipschitz constant $(1+\frac 1 C)$ is close enough to $1$.
Integrating in polar coordinates over the  ball $B^x\subset (\R^n, |\cdot|_x)$ we infer:
$$u(x)\cdot \mathcal L^n (B^x\setminus B_x) \leq r^{n-1} \cdot \Big (\int _{v\in S^{n-1} _x} ( 2\cdot r \cdot \int _0 ^r   K(x+sv) \, ds)
\cdot d\mathcal H^{n-1} \Big ) =$$
$$=2r^n \cdot \int _{B^x}  K(y) \cdot |y-x|^{n-1} \cdot u(x) \cdot d\mathcal L^n  \;,$$
where $S^{n-1} _x$ is the unit sphere in $(\R^n, |\cdot|_x)$.


%Suppose $(1-\eps)g_0\le g\le (1+\eps)g_0$.  Then for any $x$ and $v$ with $||v||_x=1$ we it holds that  If $\||v\|_0 -1| \leq \eps$.
 %Set $K_v (x) = D_x g(v,v) (v)$.   If $\||v\|_0 -1| \leq \eps$  then
%$|D_x||v||_x(v)|\le |K_v (x)| $.

 %$\lim _{t\to 0} \frac {\||v\|_{x+tv} - \|v\| _x|} t \leq |K_v (x)|$.

 %Set $K(x)= \Sigma |K_{v_i}(x)|$, where $v_i$ runs through a set of $n^3$ unit vectors in general position.
 %Then $K_v(x) \leq K (x)$, for all unit (with respect to   $g_x$ for any $x$ ) vectors $v$.

 %Set $B^x = \{ y | \|y-x\| _{g_x} \leq r \}$ and $B_x = B_r (x)$.  Let $\vol ^x$ be the constant
 %volume form coresponding to the constant flat metric $g_x$.


  % The volume of $B^x$ with respect to $\vol ^x$ is equal to $\omega_n r^n$. Thus we only need to estimate
  %the volume of $C_x$ and the difference between $\vol ^x$ and $\vol$.

To get a  similar estimate of the other summand in  \eqref{eq:summand},
we only need to  recall  the following inequality from \cite[Lemma 4.1]{Evans}, valid for any $\mathcal C^1$ function $u$ on a Euclidean ball
\[
 \int _{|x-y| < r} |u(y)-u(x)| \cdot d\mathcal L^n (y) \leq  C_2\cdot r^{n}\cdot  \int _{|x-y| < r}|u'(y)| \cdot |y-x| ^{1-n} \cdot d\mathcal L^n (y) \; .
 \]
Taking both estimates together with \eqref{eq:Kg},  embedding $B_x$ and $B^x$ in slightly larger Euclidean balls  and using  that $\frac 1 2 < u <2$, we conclude:

\[
\omega _n \cdot r^n - b_r (x) \leq C_3\cdot r^n \cdot \int _{|x-y|< \frac 4 3 r}  |g'(y)| \cdot |y-x|^{1-n} \cdot d\mathcal L^n \; .
\]

We divide both sides  by $\omega _n \cdot  r^n$ and integrate over $A$. Using that the biLipschitz constant is close enough to $1$,  we see:
$$v_r (A) \leq C_4 \int _A  \Big (\int _{|x-y|< \frac 4 3 r}  |g'(y)|  \cdot |y-x|^{1-n}\cdot d\mathcal L^n (y) \Big ) \cdot d\mathcal L^n (x)  \leq$$
$$\leq C_4 \int _{B_{2r} (A)} \Big ( \int  _{|x-y|< \frac 4 3 r} |g'(y)|  \cdot |y-x|^{1-n} \cdot d\mathcal L^n (x) \Big ) \cdot d\mathcal L^n (y) = $$
$$= \frac 4 3 C_4 \int _{B_{2r} (A)} |g'(y)| \cdot r \cdot d\mathcal L^n (y) \; ,$$
where we have used \lref{lem:exchange} in the second inequality.
   This  finishes the proof of  Proposition~\ref{prop-smooth}.
\end{proof}


%\begin{rem}
%Estimating  $v_r (A)$ from below seems to be more difficult.  We do not need a lower estimate  in the general case in this paper because in our applications $v_r$ is nonnegative due to Bishop--Gromov volume comparison. \footnote{ {V.:\color{red} do we need to go into this?} A.: No it is not needed. Should we mention the problem somewhere or just delete it?  }
%\end{rem}

\subsection{Functions of bounded variations}
%before proceeding let us briefly recall some known facts about $DC$ functions and $DC$ maps.
Let $U$ be an open subset of $\R^n$. A function $f\in L^1 (U)$ is of class  BV (bounded variation) if its first partial derivatives,
 $\frac{\partial f}{\partial x^i}$ (here and below always in the sense of distributions) are signed Radon measures with finite mass
 $|\frac{\partial f}{\partial x^i}|(U)$. We denote by $[Df ]$ the Radon measure $\sum_i |\frac{\partial f}{\partial x^i}|$ on $U$.  If $f:U\to \R$ is a BV function, which is continuous on a subset $R \subset U$ with $\mathcal H^{n-1} (U\setminus R)=0$ then the Radon measure $[Df]$ vanishes on all Borel subsets $A\subset U$ with $\mathcal H^{n-1} (A)<\infty$, \cite{Goffmann}.

Let $f:U\to \R$ be of class BV.  Then  for $\mathcal H^n$-almost every point
$x\in U$ there exists an affine  function $\hat f _x:\R^n\to \R$, such that for the BV function $h_x=f-\hat f_x$  we have
\begin{equation}\label{DC-ae-diff}
\lim_{r\to 0}\frac 1 {r^{n+1}}  \int _{B_r(x)}|h_x| =0 \; ;\; \;  \text{and} \;  \; \;    \lim _{r\to 0}  \frac 1 {r^n} \cdot   [Dh_x] (B_r (x)) =0 \;
\end{equation}
see \cite[ Theorem 6.1 (2),(3)]{Evans}  for the second and   the H\"older inequality and  \cite[ Theorem 6.1 (1)]{Evans}  for the first inequality.
%A function $f\co U\to \R$ is called $DC$ if it can be locally represented as  $h-g$ where $h$ and $g$ are  semiconcave.
%We refer to \cite{Per-DC} and \cite{Shioya} for more details on the results  about DC-functions used below.

%A map $f=(f_1,\ldots, f_m)\co U\to R^m$ is called $DC$ if each of its coordinates is $DC$.
 %$DC$ functions form an algebra and a composition of two $DC$ maps is $DC$. If $f,g$ are $DC$ functions and $g$ is never $0$ on $U$ then $\frac f g$ is $DC$.
%Partial derivatives of $DC$ functions, here and below always in the sense of distributions, are  functions of bounded variations, abbreviated as $BV$-functions below.
%Any partial derivative of a function of bounded variation is a signed Radon measure. Therefore (distributional) second partial derivatives of $DC$ functions are signed Radon measures.

 %Just like semiconcave functions, $DC$ functions have second differential  a.e. and their partial derivatives which exist a.e. are differentiable a.e.

\subsection{Almost Riemannian metric spaces}
The following definition provides a suitable description of a  large part of any Alexandrov space, see Section~\ref{sec:Alex}.

 Let  $C =C(n)$ be the constant determined in  Proposition \ref{prop-smooth}. We will call  a locally geodesic metric space
 $X$ an \emph{almost Riemannian metric space}  if it has the following properties.
 \begin{enumerate}
 %\item  $X$ is an $n$-dimensional Lipschitz manifold.
 \item  There is   a Borel subset $R\subset X$, called the subset of \emph{regular points} with $\mathcal H^{n-1} (X\setminus R)=0$.
 \item Any   minimizing geodesic $\gamma$ in $X$ can be approximated by curves $\gamma _i$ in $R$, such that the lengths of $\gamma _i$ converge to the length of  $\gamma$.
\item For any  $x\in X$, there is a neighborhood $U$ of $x$, called a \emph{regular chart}, and  a biLipschitz map
$\phi:U\to O$ onto an open subset $O\subset \R^n$, with the biLipschitz constant less than $(1+\frac 1 C)$.
 \item There a continuous Riemannian tensor $g_{ij} $ on $\phi (U\cap R)$ such that $g_{ij}$ is a  function of bounded variation on $O$
for each $1\leq i,j \leq n$.
 \item The length of any curve $\gamma \subset R$ can be computed as the length of $\phi (\gamma )$ via this Riemannian tensor $g$.
% \item The metric on  $X$ can be reconstructed from the restriction to $R$. \footnote{I do not now how to express this in a good way.}.
\end{enumerate}

For any regular chart $U$ as above, we set $\mathcal N_0 (U)$ to be the Radon measure $[g']$ on $U$ given as the sum of
the Radon measures $[D g_{ij}]$ over the coordinates $ g_{ij}$ of the metric tensor $g$.
For an almost Riemannian metric space $X$, we define an outer measure $\mathcal N$ on $X$  in the following way.
For a subset $A\subset X$, we consider all  coverings $A\subset \cup U_i$ by  countably many regular charts $U_i$ and let $\mathcal N(A)$
to be the infimum of the sums $\sum_i \mathcal N_0 (U_i)$  over all such coverings.  This is indeed an outer measure, which takes finite values on compact subsets.  Since $\mathcal N$  satisfies the Caratheodory criterion, \cite[Theorem 1.9]{Evans},  it is indeed a Radon measure.  We will call $\mathcal N$ the \emph{minimal metric derivative measure} on the almost Riemannian metric space $X$.


%We observe:
\begin{lem} \label{lem:minderiv}
Let $X^n$ be a almost Riemannian metric space and let $\mathcal N$ be its minimal  metric derivative measure.
Then $\mathcal N (A)=0$ for any Borel subset $A\subset X$ with $\mathcal H^{n-1} (A)<\infty$.
There exists a Borel subset $C\subset X$ of full $\mathcal H^n$-measure in $X$ with
$\mathcal N(C)=0$, thus $\mathcal N$ is absolutely singular with respect to $\mathcal H^n$.
\end{lem}


\begin{proof}
Clearly, both claims are local. Hence we need to verify them only in a regular chart $U$, which we identify with its image $\phi (U) \subset \R^n$.   The first statement
follows directly from the continuity of the metric tensor $g$ on the subset $U\cap R$ and the result of \cite{Goffmann}
cited above.

 In order to verify the second claim we only need to show the following statement, cf. \cite[Section 1.6]{Evans}.   For almost all $x\in U$ there is another
 regular chart $x\in V$, such that the derivative measure $[h']$ of  the Riemannian tensor $h$ in this  chart $V$, has $n$-dimensional density $0$
 at $x$, thus
 \begin{equation} \label{eq:deriv}
 \lim _{r\to 0}  \frac 1 {r^n} \cdot  {[h'] (B_r (x))}  =0 \;.
 \end{equation}
 Here and below, the ball $B_r (x)$ over which we integrate can be equally considered with respect to the Euclidean or to the original metric on $U$, since both are biLipschitz euqivalent.
 In order to prove \eqref{eq:deriv},   we follow \cite[Section 4.2]{Per-DC} and consider the Riemannian tensor $g$ of the original chart $U$.  Applying  ~\eqref{DC-ae-diff}  to  the coordinates of $g$, we find for $\mathcal H^n$-almost all
 $x\in U$  a  smooth symmetric $2$-tensor  $\hat g =\hat g_x$ on  $U$ such that  for
 $u=g-\hat g$  we have:
 \begin{equation} \label{eq:u}
 \lim_{r\to 0}\frac 1 {r^{n+1}}  \int _{B_r(x)} ||u|| =0 \; ;\; \;  \text{and} \;  \; \;
  \lim _{r\to 0}  \frac 1 {r^n} \cdot   [Du] (B_r (x)) =0 \; .
  \end{equation}
The first statement implies that $\hat g$ is indeed a Riemannian metric in a neighborhood $U_0$ of $x$.


Fix such a point $x$,  neighborhood $U_0$ and $\hat g$.
Consider a small neighborhood $W$ of $0$ in $\R^n$ and let $\xi :W\to U$ be the exponential map with respect to the metric
$\hat g$. Then $\xi (0)=x$, $D\xi  (0) =Id$  and  the  pull-back Riemannian metric $\hat h= \xi^{\ast}  (\hat g)$  has zero derivative at $0$.    Since $D\xi$ is the identity, the biLipschitz constant of the restriction $F=\xi ^{-1} \circ \phi$ to a sufficiently
small neighborhood $V$ of the point $x$ is still less than $(1 + \frac 1 C)$.      Hence, $F:V\to \R^n$ is a regular chart.

The Riemannian tensor $h$ in this chart equals $\hat h  +\xi^{\ast } (g-\hat g)$.
Now, $D\hat h (0) =0$, thus
\eqref{eq:deriv} holds for $\hat h$ instead of $h$.  For the other summand  $\xi ^{\ast} (u)$, the density estimate
\eqref{eq:deriv} follows from \eqref{eq:u} and  the fact that $\xi$ is a $C^2$-diffeomorphism
 if $W$ is sufficiently small.   This finishes the proof of \lref{lem:minderiv}.
\end{proof}


%Condition (ii) implies that the restriction of the length structure to $R$ determines the metric on $X$, at least locally.










\subsection{The upper bound on the deviation  measures}
$  $
%
 Continuing to denote by $C=C(n)$ the constant from  Proposition~\ref{prop-smooth} we show:

\begin{cor} \label{cor-dc-vr}
Let $U$ regular chart of an almost Riemannian metric space $X$. Identifying $U$ with its image $O=\phi (U)$, let
$g$ be the metric tensor and the measure $\mathcal N_0 =[Dg]$ the  derivative of the metric tensor.  For any Borel subset $A\subset U$ and any $r$
such that $B_{3r} (A)$ is relatively compact in $U$ we have

\[
v_r (A) \leq 2 \cdot C \cdot  \mathcal N_0 (B_{3r} (A))\, .
\]
\end{cor}

\begin{proof}
Consider a relatively compact open subset $ V \subset U$, which contains $B_{2r} (A)$.  Apply (coordinatewise) the standard mollifying construction
to the Riemannian tensor $g$.  For all small positive $\epsilon$, we thus obtain  smooth metrics $g_{\epsilon}$ on $V$ with the following properties.
The total derivatives $|g_{\epsilon} '|$, considered as measures, satisfy  $|g_{\epsilon} '|\leq \mathcal N_0$ on $V$, \cite[Theorem 5.3.1]{Ziemer}.   Since $g$ is pointwise   $\frac 1 C$-close to the background Euclidean inner product, the same is true for  $g_{\epsilon}$.  For all sufficiently small $\epsilon$ the $2r$-tubular neighborhood around $A$ with respect to $g_{\epsilon}$ is contained in the $3r$-tubular neighborhood around $A$ with respect to the original distance in $X$. Moreover, $g_{\epsilon}$ converges to $g$ pointwise at all points of $R$, \cite[Theorem 1.6.1]{Ziemer}.

 Denote by $d_{\epsilon}$ the distance function induced by $g_{\epsilon}$.  From the last statement and the  properties (2),(5) in the definition of an almost Riemannian metric space we deduce that
\[
\lim _{\epsilon \to 0}\, \Big (\sup _{x,y \in V, d(x,y) <r} |d_{\epsilon} (x,y) -d(x,y)|\Big )=0 \; .
\]
 Finally, the Hausdorff measures of the Riemannian metrics $g_{\epsilon}$ converge on $V$ to the Hausdorff measure of $V$ with respect to the original metric.

 Now the result follows directly  from \pref{prop-smooth} applied to the metrics $g_{\epsilon}$, by letting $\epsilon$ go to $0$.
\end{proof}



As a consequence of \cref{cor-dc-vr}, the minimal metric derivative measure bounds from above the deviations measure $v_r$
on any almost Riemannian metric space:





\begin{lem} \label{cor-mu-dc}
Let $X$ be a almost Riemannian metric space with the metric derivative measure $\mathcal N$.
Then for any compact subset $A\subset X$, there exists some $r_0>0$ such that for all
$r<r_0$ we have $$v_r (A) \leq r\cdot 2  (n+2)\cdot C\cdot \mathcal N (A)\,. $$
\end{lem}


\begin{proof}
Cover $A$ by finitely many regular charts $U_i$ such that $\sum \mathcal N_0 (U_i)$ is sufficiently close to $\mathcal N (A)$.
Since the covering dimension of $X$ is $n$, we find   a finite covering $V_j$ of $A$, which refines  the first covering but has intersection multiplicity less than $(n+2)$.  Considering each $V_j$ as a subchart of the corresponding chart $U_i$ we see that
 $$\sum \mathcal N_0 (V_j) \leq (n+2) \sum \mathcal N(A) \; .$$
Consider $r_0>0$ such that for any $x\in A$ the ball $B_{4r_0} (x)$ is contained in one of the sets $V_j$.   Denote by $A_j$ the set of all such $x$.
 Then  $v_r (A) \leq \sum v_r (A_j)$ and, due to \cref{cor-dc-vr},  $v_r (A_j) \leq  2 \cdot C\cdot N_0 (V_j)$.
   Combining these  inequalities finishes the proof.
\end{proof}















\section{Alexandrov spaces} \label{sec:Alex}
\subsection{Strained points}  \label{subsec:strainer}
  Strainers and strainer maps are basic tools for Alexandrov spaces, cf. \cite{BGP},  \cite{Otsu-Shioya}, \cite{Shioya} and will play an important role in the proof of \tref{alexandrovthm} below. Rather than recalling the definition, we will just list main properties of the subsets of strained points.  We fix a natural number $n$.
% and only consider Alexandrov spaces of curvature $\geq -1$ below.
There exist a positive number $A=A(n)$ such that for   all $0<r,\delta \leq  \frac 1 {A^2}$ and all $n$-dimensional Alexandrov space $X$ of curvature $\geq -1$ the following holds true.
\begin{enumerate}

\item The set $X_{r,\delta}$ of points in $X$ which have an $Ar$-long $(n,\delta)$-strainer is   open in $X$.  For $s<r$, we have $X_{r,\delta} \subset X_{s,\delta}$,  \cite[9.7]{BGP}.

\item Let a sequence    $(X_i ^n,x_i)$ of Alexandrov spaces  of curvature $\geq -1$ converge in the pointed Gromov--Hausdroff topology to an $n$-dimensional Alexandrov space $(X,x)$.
If $x\in X_{r,\delta}$ then, for all large $i$, the point $x_i$ has an $Ar$-long  $(n, \delta)$-strainer in $X_i$.

 \item Rescaling $X$ with a constant $\lambda \geq 1$ sends the subset $X_{r,\delta}$ to a subset of $(\lambda X )_{\lambda r, \delta} $ of the rescaled Alexandrov space $\lambda X$.

\item The union $X_{\delta}:= \cup _{r>0}  X_{r,\delta}$ contains the set $X_{reg}$ of all regular points of $X$. The Hausdorff dimension of the set $X\setminus (X_{\delta} \cup \partial X)$ is at most $n-2$~\cite[10.6, 10.6.1, 12.8]{BGP}.

\item    For any point $x\in X_{r,\delta}$ there are natural distance coordinates $\phi: B_{3r} (x) \to \R^n$ which are $(1+\epsilon)$-biLipschitz onto an open subset $O \subset \R^n$.  Here, $\epsilon\to 0$ as $A\to\infty$, \cite[9.4]{BGP}.

\item The natural coordinates $\phi$ can be smoothed out to satisfy the following property, \cite[Theorem B]{Otsu-Shioya}.  There exists a continuous Riemannian metric  $g$ on $\phi (X_{reg} \cap B_{3r} (x)) \subset O$  such  that for any curve $\gamma \subset X_{reg} \cap B_{3r} $ its length
  coincides with  the length of $\phi (\gamma )$  with respect to the Riemannian metric $g$.

\item  The metric tensor $g$ on a chart $O$ defined above is of bounded variation on $O$, \cite[4.2]{Per-DC} (cf. \cite{AB15}).
    \end{enumerate}


  The last three statements in the above list together with the density and convexity of the set $X_{reg}$ of regular points imply:
  \begin{cor}
  In the above notations, the subset $X_{\delta} \subset X$ is an almost Riemannian metric space, once $A$ is sufficiently large.
  \end{cor}



In fact,  the arguments in \cite[4.2]{Per-DC},  provide a slightly more precise version of (7) in the above list:

\begin{lem} \label{lem:A}
In the notations above,  the constant $A$ can be chosen  large enough, so that the following holds true.    The derivative measure $[g']$ of the Riemannian tensor $g$
in the canonical distance chart $O$ satisfies $[g'] (\hat O) \leq A \cdot  r^{n-1} $, where $\hat O$ is the image $\phi (B_{2r} (x)) \subset O=
\phi (B_{3r} (x))$.
\end{lem}

\begin{proof}
We only sketch the proof, referring to \cite{Per-DC} for details. First we fix $r=\frac 1 {A^2}$.

The fact that $g$ has bounded variation in the chart $O$  follows in \cite[Section 4.2]{Per-DC},  by writing the coordinates of $g$
as  a universal  smooth map $\Phi (f_1,....,f_{\alpha})$    of a finite number of distance functions $f_j$  on $X$  and their partial derivatives, both  expressed in the chart $\phi$.
 It is shown in \cite[Section 3]{Per-DC}, that any such distance function $f_j$  is expressed in the chart $O$ as a difference  of two  $L$-Lipschitz and $\lambda$-concave functions,
 where $L,\lambda$ depends only on the semi-concavity of the corresponding distance functions  in $X$. Since we have fixed $r>0$, these numbers $\lambda,L$ can  be chosen independently of $X$.  Thus, $f_j$ can be written in the chart $O$ as the  difference of two convex functions with universal Lipschitz constants $L'$. Therefore, for any unit vector $v\in \R^n$, we have a uniform bound on the total mass of the
 Radon measure $[\frac {\partial ^2 f_j} {\partial ^2v}] (\hat O)$.
 This implies that all partial second derivatives of $f$ have uniformly bounded mass on $\hat O$,
 cf. \cite[Theorem 6.8]{Evans}.

From this we deduce a uniform bound  $A'$ on the total mass $[g'] (\hat O)$, for the fixed value of $r_0=\frac 1 {A^2}$.


For any $r<r_0$ we rescale the space by $\frac {r_0} r$.  The total mass of the Riemannian tensor $g$ is then rescaled
by $(\frac {r_0} r)^{n-1}$.   Thus,  $$[g'] (\hat O) \leq A'\cdot (r_0 ) ^{1-n} \cdot r^{n-1} \; .$$
We finish the proof by replacing $A$ by $\max(A, A' r_0^{1-n})$.
\end{proof}





 %Let $X_{\delta}$ be the set of $\delta$-strained
 %points and $S_{\delta}$ be its complement.
%Let $p\in X_{\delta}$ and let $\{(a_i,b_i)\}_{i=1}^n$ be an $(A,\delta)$ strainer near $p$. Then the distance coordinate map
  %$x\mapsto (|x,a_1|,\ldots |x,a_n|)$ is locally $1+10\delta)$-biLipschitz near $p$ ~\cite{BGP} and according to ~\cite{Per-DC} the collection of such coordinate charts over all $p\in X_{\delta}$ gives $X_{\delta}$ a structure of a $DC^0$ manifold. Moreover, the averaged coordinate charts
  %$$x\mapsto (\oint_{B_\eps(a_1)}|xy|dvol(y),\ldots, \oint_{B_\eps(a_n)}|xy|dvol(y))$$ where $\eps\ll A$ gives $( X_{\delta},R)$ the structure of a $DC^1$-manifold for some subset of full measure $R\subset X_{\delta}$.

Now we use \cref{cor-dc-vr} to conclude:


 \begin{prop}  \label{ballmeasure}
 %The measure $\mu$ defined in the last section satisfies $\mu (X_{\delta } )< \infty$.
Let $C=C(n),A=A(n)$ be the constants from \pref{prop-smooth} and \lref{lem:A}.
For any point $x\in X_{r,\delta}$, any $s<r$ and any
 Borel subset $K\subset B_r (x)$ the deviation measure $v_s $ satisfies
  $v_s (K)  \leq 2\cdot C\cdot A \cdot r^{n-1}.$
 \end{prop}





\subsection{Decomposition in good balls}
Let the constant  $A$ be as above.
A ball $B_r(x)$ in $X^n$ will be called \emph{good} if $x\in X_{r,\delta}$.
%if any point $p\in B(x,2 R)$ is  $(AR,\delta)$-strained with respect to a fixed
%collection of points say $a_1,\dots a_n,b_1,\dots,b_n$.
%Recall that in this case
%the distance map $B(x,r)\to \RR^m$
%$$p\mapsto(|a_1p|,\dots,|a_np|)$$
%is bi-Lipschitz with bi-Lipschitz constant close to 1.
%
A ball $B_r(x)$ in $X$ will be called \emph{bad} if it is not good.
The aim of this subsection is the proof of a controlled covering result, cf. Problem \ref{qe:control}.
%Let us denote by $\alpha_m$ the volume of unit ball in the $m$-dimensional Euclidean space.




\begin{prop}\label{prop:covering}
Let $X^n$ be an $n$-dimensional Alexandrov space without boundary.  For every compact $W\subset X$, every $\alpha >n-2$   there exists
a positive number $  q =q(W,\alpha )  >0$ with the following property. For every $x\in W$ and every $s<1 $
 there exists a countable collection of good balls $B_m=B_{r_m}(x_m) \subset X$
such that
\begin{enumerate}
\item $r_m<s$ for all $m$.
\item $ \mathcal H^n \big(B_s (x) \setminus   (\cup _m B_m )\big)=0$.
\item  $\sum_m r_m^{\alpha}< q \cdot s^{\alpha}$.
\end{enumerate}
\end{prop}

The proof will be obtained by a recursive application of the following lemma.

\begin{lem}\label{lem:covering}
There is an integer $N  =N(W,\alpha) $ with the following property. For any $p\in W$ and $\rho < 1$
the ball $B_{\rho}(p)$ can be covered by at most $N$ balls
 $B_i=B_{r_i}(x_i)$ such that $r_i <\rho$, for all $i$, and
$$ \sum_{i\in \BAD}r_i^{\alpha}< \tfrac12\cdot \rho^{\alpha} \; ,$$
where $i\in \BAD$ means that $B_i$ is a bad ball.
\end{lem}

\parit{Proof.}
Assume the contrary. Thus we can find   a sequence of  balls
$K_l=B(p_l,\rho_l)$
such that $p_l\in W$,
$\rho_l<1$ and
one needs at least $l$ balls to cover $K_l$, so that the conditions in the lemma are fulfilled.

Taking a subsequence we may assume that the following limit exists in the pointed Gromov--Hausdorff metric.
$$(\tfrac1{\rho_m}\cdot X,p_m)\GHto (Y,p).$$
Since the points $p_m$ range over a compact subset of $X$ and   $\rho _m <1$, the
sequence is non-collapsing, i.e. $Y$ is an $n$-dimensional Alexandrov space.
By Perelman's stability theorem, $\partial Y$  is empty.
Therefore,  $S:= (Y\setminus Y_{\delta} ) \cap \bar B_2 (p)$ is a compact set of   Hausdorff dimension $\le n-2$.

By the definition of Hausdroff dimension, we can cover $S$ by a finite number of balls
$B_i=B_{r_i}(x_i)$ such that
$$\sum_ir_i^{\alpha} < (\tfrac12)^{\alpha}.$$


Any point in the remaining compact set $K\backslash (\cup_i B_i)$
is contained  in $Y_{\delta}$. Therefore a small ball centered at any point of this set is good.
By compactness, we can cover $K\backslash (\cup_i B_i)$ by a finite number of good balls.
Let $N$ be the total number of balls in the obtained covering of $K$.

Lifting the constructed covering to $K_l$, for all $l$ large enough,
we cover
the ball $K_l$ by at most $N$ balls satisfying the conditions of the lemma.
This  contradiction to our assumption finishes the proof of the lemma.
\qeds

Now it is easy to derive
\begin{proof}[Proof of Proposition \ref{prop:covering}]
%Covering $\bar W$ by a finite number of small balls we may assume that $W =B_{r_0} (x_0)$ such that $B_{3r_0} (x_0)$ does not intersect
%$\partial X$ and such that $r_0<1$.
Cover $B_{s} (x)$ by $N$ balls as in \lref{lem:covering}  and call this covering $\mathcal F_1$.  Now cover every bad
 ball from the covering $\mathcal F_1$ by at most $N$ balls provided by  \lref{lem:covering}. Together with the good balls from $\mathcal F_1$ the new balls define a covering $\mathcal F_2$ of $B_s (x)$. Proceeding in this way define for each natural number $k$ a covering $\mathcal F_k$ of $B_s(x)$.

 Denote by $g_l^{+}$ and $g_l ^-$ the sum of $r_i ^{\alpha}$ over good, respectively bad balls $B_{r_i} (x_i)$ in the covering $\mathcal F_l$.
 Then, by construction, $g_{l+1} ^- < \frac 1 2 g_l ^-$ and $g_{l+1} ^+ \leq g_l ^+ + N\cdot g_l ^-$. Therefore, $g_l^- \leq  2^{-l} \cdot  g_1 ^-$
 and $g_l^+$ is uniformly bounded form above.
 The volume of the union of bad balls in $\mathcal F_l$ is at most $g_l^-$ and converges to $0$ as $l$ goes to $\infty$.

 Let  $\mathcal F$ be the set of all good balls $B_j =B_{r_j} (x_j)$ from all the  coverings $\mathcal F_l$.
   Then $\mathcal H^n (B_s (x) \setminus (\cup _{\mathcal F} B_j  )   \leq  \lim _{l\to \infty} g_l ^- =0$.  On the other hand,
   by construction, $\sum _{B_j\in \mathcal F} r_j^{\alpha} = \lim _{l\to \infty} g_l^+  \leq  3N \cdot s^{\alpha}$.
	
		Setting $q=3\cdot N$ finishes the proof.
\end{proof}





\subsection{Final step}
Now we can provide the
\begin{proof}[Proof of Theorem~\ref{alexandrovthm}]
Let $X$ be a fixed $n$-dimensional Alexandrov space.   By the inequality of Bishop--Gromov,  the deviations measures $v_r$ are uniformly bounded from below
by a quadratic term in $r$. Thus  in order to control the mm-boundary  we only need to bound $v_r$ from above on balls in $X$.

 Let the constants $A,C$ be chosen as above, so that \pref{ballmeasure} can be applied.

Let us first assume that $\partial X$  is empty.
 Let $W\subset X$ be an arbitrary  compact subset. Fix $\alpha =n-\frac 3 2$ and choose the constant $q$ as in Proposition \ref{prop:covering}.
For any $x\in W$ and $s<1$ consider the  good  balls $B_i=B_{r_i} (x_i)$ provided by Proposition \ref{prop:covering} and set $K'=\cup _i B_i$. Let $r<\frac 1 {A^2}$ be sufficiently small.  Since $\mathcal H^n (K\setminus K')=0$
we have $ v_r (K)=v_r (K')$.





For all $m$ with $r<r_m$, we apply \pref{ballmeasure} and infer
$$v_r(B_m \cap K)\le 2\cdot  C\cdot A \cdot r\cdot r_m^{n-1} $$
On the other hand, for $r_m<r$, we have
$$v_r (B_m\cap K) \leq \mathcal H^n  (B_m) \leq 2 \cdot \omega _n  \cdot r_m ^n < 2\cdot \omega_n \cdot r \cdot r_m ^{n-1} $$
Summing up  and using $r_m^{n-1} < r_m ^{\alpha}$ we obtain
\begin{equation} \label{eq:density}
v_r (K) \leq \sum_m v_r (B_m \cap K) \leq (2\cdot C\cdot A+2\cdot \omega _n) \cdot q\cdot  r \cdot s^{\alpha} \; .
\end{equation}

This proves that $X$ has locally finite mm-boundary.  As already mentioned and used above, any  signed Radon measure $\nu$ obtained as a limit of a sequence $v_{r_j} /r_j$ for some $r_j\to 0$ must be non-negative, hence a Radon measure.    We fix such $\nu$.

Inequality \eqref{eq:density} implies that $\nu$ has finite $\alpha$-dimensional density at every point of $X$, in particular, $\nu$  vanishes on subsets of Hausdorff dimension $\leq n-2$.
  Thus $\nu (X \setminus X_{\delta } )=0$.

Recall that $X_{\delta}$ is an almost Riemannian space. Denote, by $\mathcal N $ its minimal metric derivative measure. We extend it to a measure on all of $X$ (still denoted by $\mathcal N $) by setting it to be $0$ on $X\setminus X_{\delta}$.
By \lref{cor-mu-dc} the Radon measure  $\nu$  is absolutely continuous with respect to  $\mathcal N $ on compact subsets of $X_{\delta}$.
  Now (1) and (3) of \tref{alexandrovthm} follow from \lref{lem:minderiv}.
 This finishes the proof in the case $\partial X=\emptyset$.



%since  $$\sum_{r_n< r} \vol_m W'_n\le \alpha_m\cdot \sum_{r_n< r}r_n^m.$$

Assume now that $\partial X\ne\emptyset$ and consider the doubling $Y=X\cup _{\partial X} X$ of $X$. Consider $X$ as a convex subset of $Y$ and let $K\subset X$ be compact.  We find a constant $L>0$ such that for all sufficiently small $r>0$, we have  $\mathcal H^n (K\cap B_{2r} (\partial X)) \leq L \cdot r$. (This follows, for example, by the coarea formula using the Lipschitz properties of the gradient flow of the distance function $d(\cdot, \partial X)$ which is semiconcave).   On the other hand, for $x\in K\setminus B_r (\partial X)$, the volumes of the $r$-ball in $X$ and in $Y$ coincide.  Using that $Y$ has locally finite mm-boundary, we deduce that $v_r(K)$  (computed in the space $X$) is bounded from above  $L\cdot r + \tilde v_r (K)$, where $\tilde v_r (K)$ is the deviation measure of $K$ considered as a subset of $Y$.  This implies that $v_r /r$ is uniformly bounded for $r\to 0$.    Thus $X$ has locally finite mm-boundary as well.

Any limit of a sequence $v_{r_j} /r_j$ for some $r_j\to 0$ must be  again non-negative, hence a Radon measure.  Outside of $\partial X$
   $\nu$ coincides  with the restriction of the corresponding measure defined on $Y$.   From the corresponding statement about  $Y$ we deduce that  $\nu$ is absolutely singular with respect to $\mathcal H^n$. Moreover, $\nu$ vanishes on subsets $S\subset X\setminus \partial X$ with finite $\mathcal H^{n-1}  (S)$.

   It remains to prove (2), i.e. to show that  the restriction of $\nu$ onto $\partial X$ is at least
   $c\cdot \mathcal H^{n-1}$ for  universal constant $c=c(n)$. This statement is local on $\partial X$ and needs to be verified only in small neighborhoods of points $x$  whose tangent $T_xX$
   are isometric to flat halfspaces.

	We fix such a point $x\in \partial X$. We further fix a sufficiently small $\epsilon >0$ and find
	a small neighborhood $U$ of $x$ in $X$ which is
   $(1+\epsilon)$-biLipschitz to a half-ball in the Euclidean space.  Choose an arbitrary $s>0$ such that $B_{2s} (x)\subset U$.
	Let $K= \bar B_s (x) \cap \partial X$ be the closed ball of radius $s$ in $\partial X$  with respect to the ambient metric.  Due to \cite[Section 1.6]{Evans},
	 it is sufficient to prove that $\nu (K) \geq  c_0 \cdot s^{n-1}$  for a universal constant $c_0$ depending only on the dimension.
	
	In order to prove this inequality, we consider any open neighborhood $V$ of $K$ in $X$.   For all small $r>0$, the neighborhood $V$ contains $B_{2r} (K)$.  Once $\epsilon$ has been chosen small enough,   the ball   $B_r (z)$ in $X$ has  volume at most
   $ (1-k_1)  \cdot \omega _n \cdot  r^n$,  for
	any point $z\in  B_{\frac 1 {10} r} (K)$. Here  $ k_1=k_1(n)>0 $ is a universal constant.
	 Moreover,  the set  $B_{\frac 1 {10} r} (K)$  has volume at least $\frac 1 {20} \cdot r\cdot \cdot \omega _{n-1} \cdot   s^{n-1}$.
	Integrating over $V$  (and using the inequality of Bishop--Gromov on the complement of   $B_{\frac 1 {10} r} (K)$) we deduce:
	$$v_r (V) \geq k_1\cdot \frac 1 {20} \cdot \omega _{n-1}  \cdot r\cdot s^{n-1}  -  k_3 \cdot r^2 \;,$$
	for some $k_3$ depending only on the volume of $V$ and independent of $r$.
	Dividing by $r$ and letting it go to $0$ we obtain $\nu (V) \geq k_4\cdot s^{n-1}$, for a universal constant $k_4>0$.
 Since the neighborhood $V$ of $K$ was arbitrary,
	we infer the same inequality for $K$ instead of $V$, finishing the proof.
\end{proof}














\section{Questions and Comments} \label{sec:final}
\subsection{Manifolds}
The notions of mm-boundary and mm-curvature are very easy to define but   difficult to control even in  simple situations.
For instance, the examples mentioned in the introduction require some amount of computations and estimates.  On the other hand, constructing examples in which the mm-boundary is not locally finite  also seems to be highly non-trivial.  The first  question in this direction is:


\begin{quest}
 Construct a compact manifold with a continuous Riemannian metric
that does not have finite mm-boundary.
\end{quest}

 The following problem is motivated by our approach to \tref{alexandrovthm} in Sections \ref{sec-BV-estimate}, \ref{sec:Alex}.
%We have reduced the question of triviality of mm-boundary in Alexandrov spaces to a question
%about the existence of good coordinates for almost Riemannian spaces.

\begin{quest} \label{qe:min}
Let $X$ be an almost Riemannian space. Can  the minimal metric derivative measure be non-zero?
\end{quest}

 Note that the minimal metric derivative measure must vanish if the metric can be locally defined  by a Riemannian
tensor of class $W^{1,1}$,  since the metric derivative measure is absolutely singular with respect to the Hausdorff measure by \lref{lem:minderiv}.




The following question is motivated by  \lref{cor-mu-dc}   and potential  applications to geodesic flows of spaces with curvature bounded from above, cf. Problem \ref{qe:CAT}.

\begin{quest}  \label{qe:BV}
Let $X$ be an  almost Riemannian space. Can one use the minimal metric derivative measure in order  to control the deviation measures $v_r$ from below?
\end{quest}

\subsection{Surfaces and  hypersurfaces}
The answer to the following question is not trivial in view of  Example \ref{ex:cone}.


\begin{quest}
Can one express the mm-curvature of an Alexandrov  surface in terms of its curvature measure?
\end{quest}

In view of \tref{intsurface} it is reasonable  to expect  an affirmative answer to the following question
\begin{quest}
Do  convex hypersurfaces of $\R^n$ have locally finite mm-curvature?
\end{quest}


 A natural approach to this question is related to  the following conjectural generalization of a theorem of Bonk and Lang~\cite{Bonk-Lang}:
 \begin{quest}
 Let $X$ be a convex hypersurface sufficiently close to a flat  hyperplane. Can we bound the optimal biLipschitz constant for maps into the Euclidean space in terms of the total scalar curvature?
 \end{quest}

Some natural generalizations of our \tref{thmconv} are possible. Probably, slightly refined arguments can be used to
prove that any DC-submanifold of a Euclidean space has trivial mm-boundary. Using the embedding theorem of Nash, this would also provide an easy generalization of \tref{thmconv} and \tref{thmfirst} to convex hypersurfaces of smooth Riemannian manifolds.



\subsection{Alexandrov geometry and beyond}
As the next  generalization of \tref{thmfirst}, one should study  the case of smoothable Alexandrov spaces.

\begin{quest}
Does the mm-boundary vanish  in smoothable Alexandrov spaces?   Are there relations to scalar curvature measures defined in \cite{LP}?
\end{quest}


 Due to the observation after Problem \ref{qe:min}, the triviality of mm-boundary would follow from the existence of slightly smoother coordinates than the ones provided by Perelman's DC-structure.
\begin{quest}
Let $X$ be an Alexandrov space. Can one introduce coordinates on a neighborhood of the set of regular points, such that the metric is locally given by a Riemannian tensor of class $W^{1,1}$?
\end{quest}
In the two-dimensional case, the answer to this question is ``yes" by the work of Reshetnyak.


Due to  \tref{thmmain}, an affirmative answer to the following question should be expected:
\begin{quest}
Are there connections between the size of the mm-boundary of an Alexandrov space $X$ and the "average size" of the cut loci of  points in $X$.
\end{quest}



Should one have a chance to go beyond mm-boundary and towards mm-curvature, one would definitely need to improve the
decomposition statement \pref{prop:covering}, which provides a good geometric control of the size of the set of singular points of an Alexandrov space.
% Note the proof works for any number $\alpha >n-2$. However, the optimal estimate should be the boundary case:
\begin{quest} \label{qe:control}
Can one replace $\alpha>n-2$ by $\alpha=n-2$ in the statement of \pref{prop:covering}?
\end{quest}


% In the proofs of our  main results about convex hypersurfaces and Alexandrov surfaces we deduced local  volume bounds from
%a much stronger bound on the optimal  local  biLipschitz constant (for maps into the Euclidean space).
It is interesting to understand if our results provide a quantitative version of biLipschitz closedness of small balls to Euclidean balls. It is known~\cite{BGP} that there exists $\kappa(n,\delta)\to 0$ as $\delta\to 0$ such that if  $X=X^n$  is an Alexandrov space of curvature $\ge -1$, $x\in X$  such that  $\omega _n r^n -\mathcal H^n (B_r (x)) \leq \delta  \cdot r ^n$ then $B_{\frac r 4} (x)$ is  $(1+\kappa(n,\delta))$-biLipschitz  to a Euclidean ball.
\begin{quest}
Can  $\kappa(n,\delta)$ above be chosen 
%to be linear in $\delta$, i.e. to be
 of the form $C(n)\cdot \delta$? 
%More precisely:}
%Does there exist a constant $C=C(n)$  such that for any Alexandrov space $X=X^n$ {\color{red} of curvature $\ge -1$}, any $x\in X$ and any $\delta <\frac 1 C$ the following holds true.
%If $\omega _n r^n -\mathcal H^n (B_r (x)) \leq \delta  \cdot r ^n$, then
%the ball $B_{\frac r 4} (x)$ is  $(1+C \cdot \delta )$-biLipschitz  to a Euclidean ball.
\end{quest}





 It is natural to look at what happens for  curvature bounded above:


\begin{quest}  \label{qe:CAT}
Can one estimate and use the mm-boundary in geodesically complete spaces with upper curvature bounds to study the geodesic flow?
\end{quest}

Finally, it seems reasonable to expect some generalizations to spaces with Ricci curvature bounds, for instance:


\begin{quest}
Can one control the mm-boundary of noncollapsed limits of Riemannian manifolds with Ricci curvature bounded below? Can one expect something like a geodesic flow in this setting?
\end{quest}

From the work of Cheeger and Naber \cite{Ch-Na-codim4} it should follow that on any non-collapsed limit of manifolds with both-sided Ricci curvature bounds, the mm-curvature is locally finite and mm-boundary is zero. Vanishing of the mm-boundary should then imply that  the geodesic flow is defined almost everywhere  and preserves the Liouville measure by the same argument as in the proof of Theorem~\ref{thmmain}.
\bibliographystyle{alpha}
%\bibliography{mmm}
%\end{document}
\begin{thebibliography}{KMS01}

\bibitem[AB15]{AB15}
L.~Ambrosio and J.~Bertrand.
\newblock {DC} {C}alculus.
\newblock arXiv:1505.04817, 2015.


\bibitem[Ale16]{Alesker}
S.~Alesker.
\newblock Some conjectures on intrinsic volumes of riemannian manifolds and
  {Alexandrov} spaces.
\newblock arXiv:1611.09546, 2016.

\bibitem[AZ67]{AZ}
A.~D. Aleksandrov and V.~A. Zalgaller.
\newblock {\em Intrinsic geometry of surfaces}.
\newblock Translated from the Russian by J. M. Danskin. Translations of
  Mathematical Monographs, Vol. 15. American Mathematical Society, Providence,
  R.I., 1967.

\bibitem[Bam16]{Bamler}
R.~Bamler.
\newblock Structure theory of singular spaces.
\newblock arXiv:1603.05236, 2016.

\bibitem[BB95]{BallmannBrin}
W.~Ballmann and M.~Brin.
\newblock Orbihedra of nonpositive curvature.
\newblock {\em Inst. Hautes \'Etudes Sci. Publ. Math.}, (82):169--209 (1996),
  1995.

\bibitem[BBI01]{BBI01}
D.~Burago, Yu. Burago, and S.~Ivanov.
\newblock {\em A course in metric geometry}, volume~33 of {\em Graduate Studies
  in Mathematics}.
\newblock American Mathematical Society, Providence, RI, 2001.

\bibitem[Ber02]{Bernig-Alex}
A.~Bernig.
\newblock Scalar curvature of definable {A}lexandrov spaces.
\newblock {\em Adv. Geom.}, 2(1):29--55, 2002.

\bibitem[Ber03]{Bernig-CAT}
A.~Bernig.
\newblock Scalar curvature of definable {CAT}-spaces.
\newblock {\em Adv. Geom.}, 3(1):23--43, 2003.

\bibitem[BGP92]{BGP}
Yu. Burago, M.~Gromov, and G.~Perelmann.
\newblock A.{D}.{Alexandrov} spaces with curvatures bounded below.
\newblock {\em Russian Math. Surveys}, 47(2):1--58, 1992.

\bibitem[BL03]{Bonk-Lang}
M.~Bonk and U.~Lang.
\newblock Bi-{L}ipschitz parameterization of surfaces.
\newblock {\em Math. Ann.}, 327(1):135--169, 2003.

\bibitem[CN15]{Ch-Na-codim4}
J.~Cheeger and A.~Naber.
\newblock Regularity of {Einstein} manifolds and the codimension 4 conjecture.
\newblock {\em Annals of Mathematics}, pages 1093--1165, November 2015.

\bibitem[EG15]{Evans}
L.~C. Evans and R.~F. Gariepy.
\newblock {\em Measure theory and fine properties of functions}.
\newblock Textbooks in Mathematics. CRC Press, Boca Raton, FL, revised edition,
  2015.

\bibitem[Fed59]{Fedcurvature}
H.~Federer.
\newblock Curvature measures.
\newblock {\em Trans. Amer. Math. Soc.}, 93:418--491, 1959.

\bibitem[Fed69]{Federer}
H.~Federer.
\newblock {\em Geometric measure theory}.
\newblock Grund. math. Wiss., Volume 153. Springer-Verlag, 1969.

\bibitem[GL80]{Goffmann}
C.~Goffmann and F.~Liu.
\newblock Derivative measures.
\newblock {\em Proc. Amer. Math. Soc.}, 78:218--220, 1980.

\bibitem[KMS01]{Shioya}
K.~Kuwae, Y.~Machigashira, and T.~Shioya.
\newblock Sobolev spaces, {L}aplacian, and heat kernel on {A}lexandrov spaces.
\newblock {\em Math. Z.}, 238(2):269--316, 2001.

\bibitem[LP17]{LP}
N.~Lebedeva and A.~Petrunin.
\newblock Curvature tensors on {Alexandrov} spaces.
\newblock {\em In preparation}, 2017.

\bibitem[OS94]{Otsu-Shioya}
Y.~Otsu and T.~Shioya.
\newblock The {R}iemannian structure of {A}lexandrov spaces.
\newblock {\em J. Differential Geom.}, 39(3):629--658, 1994.

\bibitem[Per91]{P2}
G.~Perelman.
\newblock {A. D. Alexandrov} spaces with curvature bounded below {II}.
\newblock {\em preprint}, 1991.

\bibitem[Per95]{Per-DC}
G.~Perelman.
\newblock {DC} structure on {A}lexandrov space with curvature bounded below.
\newblock preprint, http://www.math.psu.edu/petrunin/papers/papers.html, 1995.

\bibitem[Pet98]{Petparallel}
A.~Petrunin.
\newblock Parallel transportation for {A}lexandrov space with curvature bounded
  below.
\newblock {\em Geom. Funct. Anal.}, 8(1):123--148, 1998.

\bibitem[Pet07]{Petsemi}
A.~Petrunin.
\newblock Semiconcave functions in {A}lexandrov's geometry.
\newblock In {\em Surveys in differential geometry. {V}ol. {XI}}, volume~11 of
  {\em Surv. Differ. Geom.}, pages 137--201. Int. Press, Somerville, MA, 2007.

\bibitem[PP96]{PP}
G.~Perelman and A.~Petrunin.
\newblock Quasigeodesics and gradient curves in alexandrov spaces.
\newblock preprint, http://www.math.psu.edu/petrunin/papers/papers.html, 1996.

\bibitem[Res93]{Reshetnyak-GeomIV}
Yu.~G. Reshetnyak.
\newblock Two-dimensional manifolds of bounded curvature.
\newblock In {\em Geometry, {IV}}, volume~70 of {\em Encyclopaedia Math. Sci.},
  pages 3--163, 245--250. Springer, Berlin, 1993.

\bibitem[Sch93]{Schneider}
R.~Schneider.
\newblock Convex surfaces, curvature and surface area measures.
\newblock In {\em Handbook of convex geometry, {V}ol.\ {A}, {B}}, pages
  273--299. North-Holland, Amsterdam, 1993.

\bibitem[Zam82]{Zam-inv}
T.~Zamfirescu.
\newblock Many endpoints and few interior points of geodesics.
\newblock {\em Invent. Math.}, 69(2):253--257, 1982.

\bibitem[Zam92]{Zam-quest}
T.~Zamfirescu.
\newblock Long geodesics on convex surfaces.
\newblock {\em Math. Ann.}, 293(1):109--114, 1992.

\bibitem[Zie89]{Ziemer}
W.~P. Ziemer.
\newblock {\em Weakly differentiable functions}, volume 120 of {\em Graduate
  Texts in Mathematics}.
\newblock Springer-Verlag, New York, 1989.
\newblock Sobolev spaces and functions of bounded variation.

\end{thebibliography}




\end{document}




